---
title: "Nextflow and filtering pipeline"
output: html_notebook
---

# A. ONE-TIME PANEL FILES

Instructions to make one-time coordinate files needed for the STR calling pipeline.

## 1. Extract loci in panel from full set of shiny server data (only needs to be done once per panel)

The table of shiny server info only needs to be made once. It can then be saved on the HPC and used as input for the rest of the pipeline. Making the table will not be included as part of the pipeline.

The details of this code may need to be updated if the list of loci included in the panel is provided in a different format.

```{r}
library(dplyr)

# Import full shiny server microsatellite data table
# format: row name,chr,start,end,width,period size,motif,motif family
all_shiny <- read.delim("R:/homes/cal8150/Microsatellite-capture/Microsatellite-datasets/all_shiny.microsatellites.csv", sep = ",", header=T)

# modify row name to include prefix "MS-"
all_shiny$row.number <- paste0('MS-',all_shiny$row.number)

# import Twist file listing the locus name and the number of probes associated with that locus
status_all_probes <- read.delim("R:/homes/cal8150/Microsatellite-capture/Twist-panels/Scale up/NYU_Evrony_ScaleUp_Plus_MSI_TE-97721910_hg38/status_all_probes_NYU_Evrony_ScaleUp_Plus_MSI_TE-97721910_hg38_COUNTS.tsv")

# keep only loci targeted by at least one probe ("passing" loci)
status_passing_loci <- status_all_probes[status_all_probes$PROBE_COUNT == 1 | status_all_probes$PROBE_COUNT == 2,]

# filter full dataset to include only the "passing" loci
# filter by right joining on the locus name (include all rows found in passing loci table)
scaleup_usats_full <- right_join(all_shiny, status_passing_loci, by = c("row.number" = "TARGET"))

# drop column with number of probes
scaleup_usats_full <- scaleup_usats_full[,-9]

# Write to text file
# Final file has 1-start, fully closed coordinates
# Comma-separated file format:
# row.number,seqnames,start,end,width,period.size,motif,motif.family
output.file <- file("R:/homes/cal8150/Microsatellite-capture/Twist-panels/Scale up/scaleup_usats_full.csv","wb")

write.table(scaleup_usats_full, output.file, sep = ",", col.names = F, row.names = F, quote = F)

close(output.file)

```

## 2. Make region files for each program based on panel loci

### a. Make HipSTR input file using script (only needs to be done once per panel)

convert_to_HipSTR.sh
```{bash}
#!/bin/bash

# usage convert_to_HipSTR.sh [scaleup_usats_full.csv] [full path to output HipSTR region file]
# Makes region file for HipSTR

# scaleup_usats_full.csv is a csv that contains info directly from the shiny server for all loci in panel
# scaleup_usats_full.csv has 1-start, fully closed coordinates
# requires scaleup_usats_full.csv to be in the following order:
# row.number,seqnames,start,end,width,period.size,motif,motif.family

awk -F',' -e '{print $2"\t"$3"\t"$4"\t"$6"\t"$5/$6"\t"$1}' $1 > $2

# Output tab-delimited format: [chr] [start] [end] [period size] [number of repeats] [name]
# Coordinates are 1-start, fully-closed
```

```{bash}
# Command line input
/scratch/cal8150/uSatAnalysis/scripts/convert_to_HipSTR.sh /scratch/cal8150/uSatAnalysis/scaleup_usats_full.csv /scratch/cal8150/HipSTR/regions/scaleup_usats_hipstr.txt
```


### b. Make EH input file (only needs to be done once per panel)

To make the input file, run convert_to_EH.sh, which will automatically run eh_to_json.r afterwards. The final output will be the JSON file.

eh_to_json.r
```{r}
#!/share/apps/r/4.2.0/gcc/bin/Rscript

# usage: Rscript eh_to_json.r [temp file from convert_to_EH.sh script] [output json]
# Script will run automatically from convert_to_EH.sh
# Input file from convert_to_EH.sh is in the following tab-delimited format:
# [LocusId] [LocusStructure] [ReferenceRegion] [VariantType]

library(dplyr)
library(jsonlite)

# Import arguments from command line (in this case, from convert_to_EH.sh)
args <- commandArgs(trailingOnly = T)

# Read bed file of loci created in convert_to_EH.sh
eh_loci <- read.delim(args[1])

# Convert bed file to JSON format
# Use "pretty = T" to get proper formatting with new lines and indents
panel_json <- toJSON(eh_loci, pretty = T)

# Write JSON file to file name given to convert_to_EH.sh
write(panel_json, args[2])
```

convert_to_EH.sh
```{bash}
#!/bin/bash

# Usage: convert_to_EH.sh [scaleup_usats_full.csv] [full path to EH json output file]

# scaleup_usats_full.csv is a csv that contains info directly from the shiny server for all loci in panel
# scaleup_usats_full.csv format: row.number,seqnames,start,end,width,period.size,motif,motif.family
# scaleup_usats_full.csv has 1-start, fully closed coordinates

# Temp file is in the following tab-delimited format:
# [LocusId] [LocusStructure] [ReferenceRegion] [VariantType]
# make temp file in current directory so it can be accessed by R
# script subtracts 1 from shiny server start coordinate to make 0-based, half open (BED format)
# Final output is JSON file for ExpansionHunter 

OUT=$(mktemp -p .)

echo -e "LocusId\tLocusStructure\tReferenceRegion\tVariantType" > ${OUT}
awk -F',' -e '{print $1"\t("$7")\*\t"$2":"($3-1)"-"$4"\tRepeat"}' $1 >> ${OUT}

# remove loci that cause ExpansionHunter to abort (no way to force program to skip these loci)
# Error: too many Ns in nearby area
sed -i '/MS-187187/d' ${OUT}
sed -i '/MS-499929/d' ${OUT}
sed -i '/MS-553432/d' ${OUT}
sed -i '/MS-710766/d' ${OUT}

# Send output to R script for conversion to JSON format
Rscript /scratch/cal8150/uSatAnalysis/scripts/eh_to_json.r ${OUT} $2

# Delete temp file when finished
rm ${OUT}
```


```{bash}
# Command line input
/scratch/cal8150/uSatAnalysis/scripts/convert_to_EH.sh /scratch/cal8150/uSatAnalysis/scaleup_usats_full.csv /scratch/cal8150/expansionhunter/scaleup_usats_eh.json
```


### c. Make GangSTR input file (only needs to be done once per panel)

convert_to_GangSTR.sh
```{bash}
#!/bin/bash

# usage convert_to_GangSTR.sh [scaleup_usats_full.csv] [full path to GangSTR input file]

# scaleup_usats_full.csv is a csv that contains info directly from the shiny server for all loci in panel
# scaleup_usats_full.csv has 1-start, fully closed coordinates
# requires scaleup_usats_full.csv to be in the following order:
# row.number,seqnames,start,end,width,period.size,motif,motif.family

awk -F',' -e '{print $2"\t"$3"\t"$4"\t"$6"\t"$7}' $1 > $2

# Output tab-delimited format: [chr] [start] [end] [period size] [motif]
# Coordinates are 1-start, fully-closed
```

```{bash}
# Command line input
/scratch/cal8150/uSatAnalysis/scripts/convert_to_GangSTR.sh /scratch/cal8150/uSatAnalysis/scaleup_usats_full.csv /scratch/cal8150/GangSTR/scaleup_usats_gangstr.txt
```


### d. Make bedtools input file (only needs to be done once per panel)

convert_to_bedtools.sh
```{bash}
#!/bin/bash

# usage convert_to_bedtools.sh [scaleup_usats_full.csv] [full path to bedtools input file]

# scaleup_usats_full.csv is a csv that contains info directly from the shiny server for all loci in panel
# scaleup_usats_full.csv has 1-start, fully closed coordinates
# requires scaleup_usats_full.csv to be in the following order:
# row.number,seqnames,start,end,width,period.size,motif,motif.family

awk -F',' -e '{print $2"\t"($3-1)"\t"$4"\t"$1}' $1 > $2

# Output tab-delimited BED file: [chr] [start] [end] [locus name]
# Coordinates are 0-start, half-open
```

```{bash}
# Command line input

/scratch/cal8150/uSatAnalysis/scripts/convert_to_bedtools.sh /scratch/cal8150/uSatAnalysis/scaleup_usats_full.csv /scratch/cal8150/uSatAnalysis/scaleup_usats_bedtools.bed
```


### e. Make Picard interval list

bed_to_interval_list.sh
```{bash}
#!/bin/bash

# SBATCH --time=100


# Usage: bed_to_interval_list.sh [scaleup_usats_full.csv] [scaleup_merged_probes.txt] [reference genome dictionary file]

# scaleup_usats_full.csv is a csv that contains info directly from the shiny server for all loci in panel
# scaleup_usats_full.csv has 1-start, fully closed coordinates
# requires scaleup_usats_full.csv to be in the following order:
# row.number,seqnames,start,end,width,period.size,motif,motif.family

# scaleup_merged_probes.txt is a txt file that contains the coordinates of all regions covered by probes
# scaleup_merged_probes.txt has 1-start, fully closed coordinates
# requires scaleup_merged_probes.txt to be in the following order:
# [chr] [start] [end] [locus name]


module load jdk/11.0.9

USAT=$(mktemp -p .)
BAIT=$(mktemp -p .)

awk -F',' -e '{print $2"\t"($3-1)"\t"$4}' $1 > ${USAT}
awk -F'\t' -e '{print $1"\t"($2-1)"\t"$3}' $2 > ${BAIT}

java -jar /scratch/projects/evronylab/bin/gatk/picard.jar BedToIntervalList \
          I=${USAT} \
          O=$(dirname $2)/$(basename $1 .csv).interval_list \
          SD=$3

java -jar /scratch/projects/evronylab/bin/gatk/picard.jar BedToIntervalList \
          I=${BAIT} \
          O=$(dirname $2)/$(basename $2 .txt).interval_list \
          SD=$3

rm ${USAT}
rm ${BAIT}

# Interval list coordinates are 0-start, half-open
```

```{bash}
# Command line input

sbatch /scratch/cal8150/uSatAnalysis/scripts/bed_to_interval_list.sh /scratch/cal8150/uSatAnalysis/scaleup_usats_full.csv /scratch/cal8150/uSatAnalysis/scaleup_merged_probes.txt /scratch/projects/evronylab/reference-files/hg38-gatk/Homo_sapiens_assembly38.dict
```


######################

NEXTFLOW PIPELINE

######################


# B. SAMPLE CSV

Sample CSV format: sample ID,sex,path/to/read1.fastq,path/to/read2.fastq,trio name
```{bash}
NA12877,M,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12877_S4_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12877_S4_L002_R2_001.fastq.gz,1
NA12889,M,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12889_S6_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12889_S6_L002_R2_001.fastq.gz,1
NA12890,F,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12890_S7_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12890_S7_L002_R2_001.fastq.gz,1
NA12878,F,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12878_S5_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12878_S5_L002_R2_001.fastq.gz,2
NA12891,M,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12891_S8_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12891_S8_L002_R2_001.fastq.gz,2
NA12892,F,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12892_S9_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12892_S9_L002_R2_001.fastq.gz,2
NA12413,M,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12413_S1_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12413_S1_L002_R2_001.fastq.gz,3
NA12414,F,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12414_S2_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12414_S2_L002_R2_001.fastq.gz,3
NA12485,M,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12485_S3_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_8-5-22/NA12485_S3_L002_R2_001.fastq.gz,3
UDP-1101,M,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-1101_S1_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-1101_S1_L002_R2_001.fastq.gz,UDP-1100
UDP-1102,F,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-1102_S2_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-1102_S2_L002_R2_001.fastq.gz,UDP-1100
UDP-1104,M,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-1104_S3_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-1104_S3_L002_R2_001.fastq.gz,UDP-1100
UDP-2101,M,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-2101_S4_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-2101_S4_L002_R2_001.fastq.gz,UDP-2100
UDP-2102,F,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-2102_S5_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-2102_S5_L002_R2_001.fastq.gz,UDP-2100
UDP-2103,F,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-2103_S6_L002_R1_001.fastq.gz,/scratch/projects/evronylab/projects/TAPESTRY/raw_data/Illumina_7-25-22/UDP-2103_S6_L002_R2_001.fastq.gz,UDP-2100
```


# C. CONFIGURATION FILE

nextflow.config
```{bash}
process.executor = 'slurm'


/*
SAMPLES
*/

// CSV with fields: sample,sex,read1,read2,trio
params.samples = '/scratch/cal8150/uSatAnalysis/coriell_udp_samples.csv'


// Reference genome
params.reference = '/scratch/projects/evronylab/reference-files/hg38-gatk/Homo_sapiens_assembly38.fasta'
params.reference_index = '/scratch/projects/evronylab/reference-files/hg38-gatk/Homo_sapiens_assembly38.fasta.fai'


/*
PANEL FILES
*/

// Panel metadata
params.panel = '/scratch/cal8150/uSatAnalysis/loci/scaleup_usats_full.csv'
// Probe coordinates in Picard format
params.picard_probes = '/scratch/cal8150/uSatAnalysis/loci/scaleup_merged_probes.interval_list'
// Microsatellite coordinates in Picard format
params.picard_usats = '/scratch/cal8150/uSatAnalysis/loci/scaleup_usats_full.interval_list'
// Microsatellite coordinates in HipSTR format
params.hipstr_usats = '/scratch/cal8150/HipSTR/scaleup_usats_hipstr.txt'
// Microsatellite coordinates in GangSTR format
params.gangstr_usats = '/scratch/cal8150/GangSTR/scaleup_usats_gangstr.txt'
// Microsatellite coordinates in Expansion Hunter format
params.eh_usats = '/scratch/cal8150/expansionhunter/scaleup_usats_eh.json'
// Microsatellite coordinates in BED format
params.bedtools_usats = '/scratch/cal8150/uSatAnalysis/loci/scaleup_usats_bedtools.bed'


/*
RESULTS
*/

// Results directory
params.results = '/scratch/cal8150/uSatAnalysis/results'
// Basename for joint HipSTR VCF and final RDS file
params.outputbasename = 'coriell_udp'

// Template directory
params.script_path = '/scratch/cal8150/uSatAnalysis/templates'


/*
PROGRAMS AND MODULES
*/

params.fastqc = 'fastqc/0.11.9'
params.bwa = 'bwa/intel/0.7.17'
params.samtools = 'samtools/intel/1.14'
params.picard = '/scratch/projects/evronylab/bin/gatk/picard.jar'
params.bcftools = 'bcftools/intel/1.14'
params.hipstr = '/scratch/projects/evronylab/bin/HipSTR/HipSTR'
params.gangstr = 'gangstr/intel/2.5.0'
params.eh = 'expansion-hunter/5.0.0'
params.bedtools = 'bedtools/intel/2.29.2'
params.gatk = 'gatk/4.2.4.1'


/*
SCRIPT PARAMETERS
*/

// Number of files to split HipSTR input file into
params.split = 50
```


# D. NEXTFLOW SCRIPT

usat_calling.nf
```{bash}
// enable DSL2 syntax
nextflow.enable.dsl=2


/*

  CHANNELS

*/

// load sample metadata and FASTQ paths
samples_ch =
Channel
  .fromPath( params.samples )
  .splitCsv()


/*

  WORKFLOW

*/

workflow {

  FASTQC( samples_ch )
  REMOVE_UMIS( samples_ch )
  ALIGN_CRAM( REMOVE_UMIS.out )
  PICARD( ALIGN_CRAM.out.cram )
  SPLIT_REGIONS()
  HIPSTR( ALIGN_CRAM.out.hipstr.toList(), SPLIT_REGIONS.out.usats.flatten(), ALIGN_CRAM.out.index.collect() )
  CONCAT_VCF( HIPSTR.out.small_vcf.collect(), HIPSTR.out.index.collect() )
  SPLIT_VCF( samples_ch, CONCAT_VCF.out.joint_vcf )
  EXPANSION_HUNTER( samples_ch.join(ALIGN_CRAM.out.cram) )
  GANGSTR( samples_ch.join(ALIGN_CRAM.out.cram) )
  BEDTOOLS_COVERAGE( ALIGN_CRAM.out.cram )
  QUERY_HIPSTR( SPLIT_VCF.out.vcf )
  QUERY_EH( EXPANSION_HUNTER.out.vcf )
  QUERY_GANGSTR( GANGSTR.out.vcf )
  JOIN_CALLS( samples_ch.join(QUERY_HIPSTR.out.table).join(QUERY_EH.out.table).join(QUERY_GANGSTR.out.table).join(BEDTOOLS_COVERAGE.out.table) )
  JOIN_SAMPLES( JOIN_CALLS.out.rds.collect() )

}


/*

  PROCESSES

*/


// QC raw sequencing data

process FASTQC {

  time '2h'
  cache 'deep'

  publishDir("${params.results}/${sampleID}_results", mode: 'copy')

  input:
    // load FASTQ files
    tuple val( sampleID ), val( sex ), path( read1 ), path( read2 ), val( trio )

  output:
    // save HTML reports
    tuple val( sampleID ), path("*_fastqc.html")


  script:
  """
  module load ${params.fastqc}

  # make results directory if it doesn't already exist
  mkdir -p \$(dirname ${params.results}/${sampleID}_results)

  # run fastqc on read 1 and read 2 FASTQ files
  fastqc ${read1} ${read2}

  """
}


// Remove UMI sequence from read name if present
// UMI in read name is not compatible with Picard MarkDuplicates

process REMOVE_UMIS {

    time '4h'
    memory '16GB'
    cpus 4
    
    input:
      // load FASTQ files
      tuple val( sampleID ), val( sex ), path( read1 ), path( read2 ), val( trio )
    
    output:
      // fastq files without UMI in read name
      tuple val( sampleID ), path( "${read1.simpleName}_noUMI.fastq.gz" ), path( "${read2.simpleName}_noUMI.fastq.gz" )
      
    shell:
    '''
    zcat !{read1} | awk '{if (NR % 4 == 1) {split($1,a,":"); print a[1]":"a[2]":"a[3]":"a[4]":"a[5]":"a[6]":"a[7]" "$2} else {print $0}}' | 
    gzip -c > $(basename !{read1} .fastq.gz)_noUMI.fastq.gz
    
    zcat !{read2} | awk '{if (NR % 4 == 1) {split($1,a,":"); print a[1]":"a[2]":"a[3]":"a[4]":"a[5]":"a[6]":"a[7]" "$2} else {print $0}}' | 
    gzip -c > $(basename !{read2} .fastq.gz)_noUMI.fastq.gz

    '''
}


// Make, sort, and index CRAM with PCR duplicates removed

process ALIGN_CRAM {

  time '10h'
  cpus 12
  memory '32 GB'

  publishDir("${params.results}/${sampleID}_results", mode: 'copy')

  input:
    // load FASTQ files
    tuple val( sampleID ), path( read1 ), path( read2 )

  output:
    // output channel for most processes
    tuple val ( sampleID ), path( "${sampleID}.cram" ), path( "${sampleID}.cram.crai" ), emit: cram
    path( "${sampleID}.markdup.metrics.txt" )

    //output channels for HipSTR
    path( "${sampleID}.cram" ), emit: hipstr
    path( "${sampleID}.cram.crai" ), emit: index

  script:
  """

  module load ${params.bwa}
  module load ${params.samtools}
  module load ${params.gatk}
  

  # Align to reference genome
  # bwa mem -R [read group] -t 8 [reference fasta] [read 1 fastq] [read 2 fastq]
  bwa mem -R "@RG\\tID:${sampleID}\\tSM:${sampleID}\\tLB:${sampleID}\\tPL:ILLUMINA" -t 8 ${params.reference} ${read1} ${read2} > ${sampleID}.bam
  
  # Remove optical duplicates
  java -jar ${params.picard} MarkDuplicates \\
    I=${sampleID}.bam \\
    O=${sampleID}.markdup.bam \\
    REMOVE_DUPLICATES=false \\
    METRICS_FILE=${sampleID}.markdup.metrics.txt \\
    OPTICAL_DUPLICATE_PIXEL_DISTANCE=2500 \\
    ASSUME_SORT_ORDER=queryname \\
    CLEAR_DT=false \\
    REMOVE_SEQUENCING_DUPLICATES=true
    
  # Unmark duplicate reads
  gatk UnmarkDuplicates \
    -I ${sampleID}.markdup.bam \
    -O ${sampleID}.unmarked.bam
  
  # Sort BAM  
  java -jar ${params.picard} SortSam \\
    I=${sampleID}.unmarked.bam \\
    O=${sampleID}.unmarked.sorted.bam \\
    SORT_ORDER=coordinate; samtools index ${sampleID}.unmarked.sorted.bam
  
  # Convert to CRAM
  samtools view -F 2304 -C -T ${params.reference} --write-index -O cram -@4 -o ${sampleID}.cram##idx##${sampleID}.cram.crai ${sampleID}.unmarked.sorted.bam
  """
}


// Check hybrid capture quality with Picard CollectHsMetrics

process PICARD {

  time '2h'

  memory '4 GB'

  publishDir("${params.results}/${sampleID}_results", mode: 'copy')

  input:
    // load CRAM
    tuple val( sampleID ), path( cram ), path( index )

  output:
    // save Picard report
    tuple val( sampleID ), path( "${sampleID}_picard_usats.txt" ), path( "${sampleID}_picard_probes.txt" )

  script:
  """
  # run Picard with microsatellites as "target" coordinates
  java -jar ${params.picard} CollectHsMetrics \\
        I=${cram} \\
        O=${sampleID}_picard_usats.txt \\
        R=${params.reference} \\
        BAIT_INTERVALS=${params.picard_probes} \\
        TARGET_INTERVALS=${params.picard_usats}

  # run Picard with probes as "target" coordinates
  java -jar ${params.picard} CollectHsMetrics \\
        I=${cram} \\
        O=${sampleID}_picard_probes.txt \\
        R=${params.reference} \\
        BAIT_INTERVALS=${params.picard_probes} \\
        TARGET_INTERVALS=${params.picard_probes}

  """
}


// Divide HipSTR input file into 50 equal parts
// Will use to submit smaller, faster jobs

process SPLIT_REGIONS {

  time '20m'

  output:
    // channel divided input files
    path( "split_regions_*" ), emit: usats

  script:
  """
  # use numerical suffixes to avoid file name overlap
  split -n l/${params.split} --numeric-suffixes --additional-suffix=.txt ${params.hipstr_usats} split_regions_

  """
}


// Run HipSTR jointly on all CRAMs using divided list of microsatellites

process HIPSTR {

  time '2h'

  input:
    // load CRAMs
    // must be just CRAMs to make comma separated list for HipSTR
    path( cram )

    // load divided input files
    path( usats )

    // load all CRAM indexes
    path( index )

  output:
    // channel for VCFs to be concatenated
    path( "array_${usats}_hipstr.vcf.gz" ), emit: small_vcf

    // channel for VCF indexes
    path( "array_${usats}_hipstr.vcf.gz.csi" ), emit: index

  script:
  """
  module load ${params.bcftools}
  
  # Do joint calling on all samples

  # Make comma separated list of CRAMs
  BAMS=\$(echo ${cram} | sed 's/ /,/g')

  # run HipSTR
  ${params.hipstr} \
      --bams \${BAMS} \
      --fasta ${params.reference} \
      --regions ${usats} \
      --str-vcf array_${usats}_hipstr.vcf.gz \
      --min-reads 20 \
      --max-str-len 150 \
      --no-rmdup \
      --output-filters; bcftools index array_${usats}_hipstr.vcf.gz

  """
}


// Combine HipSTR VCFs into one VCF containing all samples and loci

process CONCAT_VCF {

  time '3h'

  publishDir("${params.results}", mode: 'copy')

  input:
    // load all HipSTR VCFs
    // must be just VCFs to output as list in script
    path( small_vcf )

    // load all HipSTR VCF indexes
    path( index )

  output:
    // channel for combined VCF
    tuple path( "${params.outputbasename}_hipstr.vcf.gz" ), path( "${params.outputbasename}_hipstr.vcf.gz.csi" ), emit: joint_vcf

  script:
  """
  module load ${params.bcftools}

  # concatenate all HipSTR VCFs and index combined VCF
  bcftools concat --allow-overlaps ${small_vcf} | bcftools sort --output ${params.outputbasename}_hipstr.vcf.gz -Oz -;
  bcftools index ${params.outputbasename}_hipstr.vcf.gz

  """
}


// Split combined HipSTR VCF by sample

process SPLIT_VCF {

  time '2h'

  publishDir("${params.results}/${sampleID}_results", mode: 'copy')

  input:
    // load sample metadata
    tuple val( sampleID ), val( sex ), path( read1 ), path( read2 ), val( trio )

    // load combined HipSTR VCF
    tuple path( joint_vcf ), path( index )

  output:
    // channel for individual VCFs
    tuple val( sampleID ), path( "${sampleID}_hipstr.vcf.gz" ), emit: vcf

  script:
  """
  module load ${params.bcftools}

  # split combined VCF by sample ID
  bcftools view -Oz -s ${sampleID} ${joint_vcf} > ${sampleID}_hipstr.vcf.gz

  """
}


// Run Expansion Hunter on each sample individually

process EXPANSION_HUNTER {

  time '10h'

  cpus 8

  memory '48 GB'

  publishDir("${params.results}/${sampleID}_results", mode: 'copy')

  input:
    // load sample metadata and CRAM
    tuple val( sampleID ), val( sex ), path( read1 ), path( read2 ), val( trio), path( cram ), path( index )

  output:
    // channel for Expansion Hunter VCFs
    tuple val( sampleID ), path( "${sampleID}_eh.vcf.gz" ), path( "${sampleID}_eh.vcf.gz.csi" ), emit: vcf

  script:
  """

  module load ${params.eh}
  module load ${params.bcftools}

  # The JSON file from convert_to_EH.sh specifies the regions to be analyzed (called a "variant catalog" by EH)
  # Must convert M/F sex designation to male/female for Expansion Hunter

  # define sex variable to fit Expansion Hunter syntax
  if [ ${sex} == "M" ]
  then
    eh_sex="male"
  else
    eh_sex="female"
  fi

  # run Expansion Hunter
  # zip and index the output VCF
  ExpansionHunter --reads ${cram} \
      --reference ${params.reference} \
      --variant-catalog ${params.eh_usats} \
      --output-prefix ${sampleID}_eh \
      --sex \${eh_sex} \
      --threads 12; bgzip -f ${sampleID}_eh.vcf; bcftools index ${sampleID}_eh.vcf.gz

  """
}


// Run GangSTR on each sample individually

process GANGSTR {

  time '10h'

  cpus 8

  memory '48 GB'

  publishDir("${params.results}/${sampleID}_results", mode: 'copy')

  input:
    // load sample metadata and CRAM
    tuple val( sampleID ), val( sex ), path( read1 ), path( read2 ), val( trio ), path( cram ), path( index )

  output:
    // channel for GangSTR VCFs
    tuple val( sampleID ), path( "${sampleID}_gangstr.vcf.gz" ), path( "${sampleID}_gangstr.vcf.gz.csi" ), emit: vcf

  script:
  """
  module load ${params.gangstr}
  module load ${params.bcftools}

  # run GangSTR
  # zip and index the output VCF
  GangSTR \
      --bam ${cram} \
      --ref ${params.reference} \
      --regions ${params.gangstr_usats} \
      --out ${sampleID}_gangstr \
      --bam-samps ${sampleID} \
      --samp-sex ${sex} \
      --min-sample-reads 20 \
      --nonuniform \
      --frrweight 0 \
      --spanweight 0 \
      --flankweight 0; bgzip -f ${sampleID}_gangstr.vcf; bcftools index ${sampleID}_gangstr.vcf.gz
  """
}


// Run bedtools coverage on each sample to get spanning read depth

process BEDTOOLS_COVERAGE {

  time '3h'

  publishDir("${params.results}/${sampleID}_results", mode: 'copy')

  input:
    // load CRAM
    tuple val( sampleID ), path( cram ), path( index )

  output:
    // channel for bedtools file
    tuple val( sampleID ), path( "${sampleID}_bedtools.txt" ), emit: table

  script:
  """
  module load ${params.bedtools}

  # Counts number of reads fully spanning microsatellite locus

  # Make column titles
  echo -e "name\tB_depth" > ${sampleID}_bedtools.txt

  # get spanning read depth at each locus
  bedtools coverage -sorted -g ${params.reference_index} -f 1.0 -a ${params.bedtools_usats} -b ${cram} | awk -e '{print \$4"\\t"\$5}' >> ${sampleID}_bedtools.txt

  # final tab-delimited output: [name] [depth]

  """
}


/*

VCF DATA EXTRACTION
  
*/


/*
HipSTR fields: https://hipstr-tool.github.io/HipSTR/
  ID - name of locus (for joining)
  FORMAT/GT - indicates whether or not the locus has been genotyped (0/1/2 = genotyped, "." = missed)
    Since HipSTR does joint calling, the genotype number may be higher than 2 if more than two genotypes are seen amongst all     samples
  FORMAT/Q - quality of the call (from manual: "Posterior probability of unphased genotype")
  FORMAT/GB - From manual: "Base pair differences of genotype from reference"
    allele 1|allele 2 (ex: 0|0) (for genotyped alleles)
  FORMAT/DP - From manual: "Number of valid reads used for sample’s genotype"
  FORMAT/DSTUTTER - From manual: "Total number of reads with a stutter indel in the STR region"
  FORMAT/DFLANKINDEL - From manual: "Total number of reads with an indel in the regions flanking the STR"
  FORMAT/ALLREADS - gives number of spanning reads supporting a certain bp difference based on alignment (for all observed bp differences)
  FORMAT/MALLREADS - gives number of spanning reads supporting a certain bp difference based on maximum likelihood haplotype (for all observed bp differences)
  FORMAT/GLDIFF - From manual: "Difference in likelihood between the reported and next best genotypes"
    Note: HipSTR only uses spanning reads to genotype
    bp diff 1|# of supporting reads;bp diff 2|# of supporting reads (for all observed bp differences) (ex: -2|2;0|111;2|1)

Why I chose these fields:
  The GT field will be used to filter for samples with successful calls from each program.
  The GB field will be used to obtain the genotypes from each caller.
  The Q field will be used to eliminate genotype calls of low quality.
  The DP and MALLREADS fields will be used to remove genotype calls with low coverage and few supporting spanning reads.
*/


// Extract relevant fields from HipSTR VCF and save in table

process QUERY_HIPSTR {

  time '1h'

  publishDir("${params.results}/${sampleID}_results", mode: 'copy')

  input:
    // load HipSTR VCF
    tuple val( sampleID ), path( vcf )

  output:
    // channel for HipSTR data table
    tuple val( sampleID ), path( "${sampleID}_hipstr_table.txt" ), emit: table

  script:
  """
  module load ${params.bcftools}

  # Extracts fields needed for quality filtering from each program's VCF

  #make HipSTR table
  echo -e "name\tH_GT\tH_GB\tH_Q\tH_DP\tH_DSTUTTER\tH_DFLANKINDEL\tH_ALLREADS\tH_MALLREADS\tH_GLDIFF" > ${sampleID}_hipstr_table.txt
  bcftools query -f '%ID\t[%GT\t%GB\t%Q\t%DP\t%DSTUTTER\t%DFLANKINDEL\t%ALLREADS\t%MALLREADS\t%GLDIFF]\n' ${sampleID}_hipstr.vcf.gz >> ${sampleID}_hipstr_table.txt

  """
}


/*
Expansion Hunter fields: https://github.com/Illumina/ExpansionHunter and https://support-docs.illumina.com/SW/DRAGEN_v310/Content/SW/RepeatGenotyping.htm
  INFO/VARID - name of locus (for joining)
  INFO/REF - number of repeats in reference genome
  FORMAT/GT - indicates whether or not the locus has been genotyped (0/1/2 = genotyped, "." = missed)
  FORMAT/REPCN - the number of repeats in the allele (for genotyped alleles)
      allele 1/allele 2 (ex: 14/15)
  FORMAT/SO - From manual: "Type of reads that support the allele; can be SPANNING, FLANKING, or INREPEAT meaning that the reads span, flank, or are fully contained     in the repeat" (for genotyped alleles)
      allele 1/allele 2 (ex: SPANNING/SPANNING)
  FORMAT/ADSP - From manual: "Number of spanning reads consistent with the allele" (for genotyped alleles)
      allele 1/allele 2 (ex: 93/67)
  FORMAT/ADFL - From manual: "Number of flanking reads consistent with the allele" (for genotyped alleles)
    Note: "flanking" reads in EH are reads that end in the repeat on one side
    allele 1/allele 2
  FORMAT/ADIR - From manual: "Number of in-repeat reads consistent with the allele" (for genotyped alleles)
    Note: "in-repeat" reads are reads that fully consist of the repeat (the repeat is longer than the read length)
    allele 1/allele 2
  FORMAT/LC - the average coverage of the locus
    
Why I chose these fields:
  The GT field will be used to filter for samples with successful calls from each program.
  The REPCN field will be used to obtain the genotypes from each caller.bet
  The SO field will be used to eliminate ExpansionHunter calls not made with spanning reads.
  The ADSP field will be used to remove genotype calls with low coverage and few supporting spanning reads.
*/

// Extract relevant fields from Expansion Hunter VCF

process QUERY_EH {

  time '1h'

  publishDir("${params.results}/${sampleID}_results", mode: 'copy')

  input:
    // load Expansion Hunter VCF
    tuple val( sampleID ), path( vcf ), path( index )

  output:
    // channel for Expansion Hunter data table
    tuple val( sampleID ), path( "${sampleID}_eh_table.txt" ), emit: table

  script:
  """
  module load ${params.bcftools}

  # Extracts fields needed for quality filtering from each program's VCF

  #make ExpansionHunter table
  echo -e "name\tE_REF\tE_GT\tE_REPCN\tE_SO\tE_ADSP\tE_ADFL\tE_ADIR\tE_LC" > ${sampleID}_eh_table.txt
  bcftools query -f '%INFO/VARID\t%INFO/REF\t[%GT\t%REPCN\t%SO\t%ADSP\t%ADFL\t%ADIR\t%LC]\n' ${sampleID}_eh.vcf.gz >> ${sampleID}_eh_table.txt

  """
}


/*
GangSTR fields: https://github.com/gymreklab/GangSTR
  CHR and POS - chromosome and start position (for joining)
  INFO/REF - number of repeats in reference genome
  FORMAT/GT - indicates whether or not the locus has been genotyped (0/1/2 = genotyped, "." = missed)
  FORMAT/Q - quality score
  FORMAT/REPCN - From manual: "Genotype given in number of copies of the repeat motif" (for genotyped alleles)
      allele 1,allele 2 (ex: 13,14)
  FORMAT/REPCI - From manual: "95% Confidence interval for each allele based on bootstrapping"
  FORMAT/DP - From manual: "Read Depth (number of informative reads)"
  FORMAT/RC - number of reads in each class (enclosing, spanning, FRR, flanking)
    # of enclosing reads,# of spanning reads,#number of FRR reads,# of flanking reads
    Note: see GangSTR paper for read type definitions
  FORMAT/ENCLREADS - gives number of spanning reads supporting a certain repeat count (from manual: "Summary of reads in enclosing class in | separated key-value pairs. Keys are number of copies and values show number of reads with that many copies.") (for all observed bp differences)
      Note: "enclosing" reads in GangSTR are defined the same way as "spanning" reads in HipSTR and ExpansionHunter 
      repeat count 1,# of supporting reads|repeat count 2,# of supporting reads (ex: 6,2|13,39|14,38)
  FORMAT/FLNKREADS - gives number of flanking reads supporting a certain repeat count (from manual: "Summary of reads in flanking class in | separated key-value pairs. Keys are number of copies and values show number of reads with that many copies.")
      Note: "flanking" reads in GangSTR are reads in which the repeat is at the end of the read.
      repeat count 1,# of supporting reads|repeat count 2,# of supporting reads
      
Why I chose these fields:
  The GT field will be used to filter for samples with successful calls from each program.
  The REPCN field will be used to obtain the genotypes from each caller.
  The Q field will be used to eliminate genotype calls of low quality.
  The DP, ENCLREADS, and FLNKREADS fields will be used to remove genotype calls with low coverage and few supporting spanning reads.
*/

// Extract relevant fields from GangSTR VCF

process QUERY_GANGSTR {

  time '1h'

  publishDir("${params.results}/${sampleID}_results", mode: 'copy')

  input:
    tuple val( sampleID ), path( vcf ), path( index )

  output:
    tuple val( sampleID ), path( "${sampleID}_gangstr_table.txt" ), emit: table

  script:
  """
  module load ${params.bcftools}

  # Extracts fields needed for quality filtering from each program's VCF

  #make GangSTR table
  echo -e "chr\tstart\tG_REF\tG_GT\tG_Q\tG_REPCN\tG_REPCI\tG_DP\tG_RC\tG_ENCLREADS\tG_FLNKREADS" > ${sampleID}_gangstr_table.txt
  bcftools query -f '%CHROM\t%POS\t%INFO/REF\t[%GT\t%Q\t%REPCN\t%REPCI\t%DP\t%RC\t%ENCLREADS\t%FLNKREADS]\n' ${sampleID}_gangstr.vcf.gz >> ${sampleID}_gangstr_table.txt

  """
}


// Join panel metadata with data tables from HipSTR, Expansion Hunter, GangSTR, and bedtools

process JOIN_CALLS {

  time '1h'

  input:
    // load data tables from HipSTR, Expansion Hunter, GangSTR, and bedtools
    tuple val( sampleID ), val( sex ), path( read1 ), path( read2 ), val( trio ), path( hipstr ), path( expansionhunter ), path( gangstr ), path( bedtools )

  output:
    // channel for sample data frames
    path( "${sampleID}.rds" ), emit: rds

  script:
  """
  Rscript ${params.script_path}/join_str_calls_nextflow.r ${params.panel} ${sampleID} ${sex} ${trio} ${hipstr} ${expansionhunter} ${gangstr} ${bedtools}

  """
}


// Combine all sample data frames in R list structure

process JOIN_SAMPLES {

  time '1h'
  memory '8 GB'

  publishDir("${params.results}", mode: 'copy')

  input:
    // load all sample data frames
    path( RDS )

  output:
    // save final R list
    path( "${params.outputbasename}.rds" )

  script:
  """
  # arguments: [RDS files] [basename for joint list]
  # Nextflow automatically outputs all RDS files in space-separated list
  Rscript ${params.script_path}/join_str_samples_nextflow.r ${RDS} ${params.outputbasename}

  """
}
```


# E. TEMPLATE SCRIPTS

### 1. join_str_calls_nextflow.r
```{r}
#!/share/apps/r/4.2.0/gcc/bin/Rscript

# usage: Rscript join_str_calls_nextflow.r ${panel} ${sampleID} ${sex} ${trio} ${hipstr} ${expansionhunter} ${gangstr} ${bedtools}

# Panel CSV is a csv that contains info directly from the shiny server for all loci in panel
# CSV has 1-start, fully closed coordinates
# requires CSV to be in the following order:
# row.number,seqnames,start,end,width,period.size,motif,motif.family

# HipSTR can't distinguish the sex of samples when performing joint calling
# This results in 2 alleles called on chromosomes X and Y in males when only one allele is present
# From the two alleles given in the H_GB field, we will choose the one associated with the higher H_MALLREADS count as the "true" allele
# If the true allele is in the allele 2 position initially in H_GB and H_MALLREADS, H_GB_1 and H_MALLREADS_1 will be swapped with H_GB_2 and H_MALLREADS_2
# This ensures that allele 1 in the final table is always the allele with most reads on the sex chromosomes
# We will replace H_REPDIFF_2 on sex chromosomes with -999 to indicate allele 2 should not be considered for concordance


library(dplyr)
library(tidyr)

args <- commandArgs(trailingOnly = T)

# load list of panel loci and metadata
panel_loci <- as.data.frame(read.csv(args[1], header = F))
# add column names to panel loci data frame
colnames(panel_loci) <- c("name","chr","start","end","width","period.size","motif","motif.family")

# sample ID
sampleID <- args[2]

# sample sex (M or F)
sex <- args[3]

# trio ID
trio <- args[4]

# load results table for each caller and bedtools
# results table for HipSTR from QUERY_HIPSTR
HipSTR <- read.delim(args[5])
# results table for Expansion Hunter from QUERY_EH
EH <- read.delim(args[6])
# results table for GangSTR from QUERY_GANGSTR
GangSTR <- read.delim(args[7])
# table of bedtools coverage from BEDTOOLS_COVERAGE
bedtools <- read.delim(args[8])


# make function for obtaining total read depth from MALLREADS, ALLREADS, and ENCLREADS
readdepth <- function(x) {sum(suppressWarnings(as.numeric(x))[seq(0, length(x), by=2)])}



# generate data frame containing all metadata and call information for the sample
results <- 
  # add sample ID and trio ID as columns to the panel loci data frame
  panel_loci %>% mutate(sampleID = sampleID) %>% mutate(sex = sex) %>% mutate(trio = trio) %>%
  # left join caller and bedtools data files to panel loci data frame
  left_join(., HipSTR, by = "name") %>% left_join(., EH, by = "name") %>% 
  left_join(., GangSTR, by = c("chr","start")) %>% left_join(., bedtools, by = c("name")) %>%
  # replace any EH genotype that has B_depth of 0 with "./." on autosomes and "." on sex chromosomes
  mutate(E_GT = case_when(B_depth == 0 & sex == "M" & chr != "chrX" & chr != "chrY" ~ "./.",
                          B_depth == 0 & sex == "F" ~ "./.",
                          B_depth == 0 & sex == "M" & (chr == "chrX" | chr == "chrY") ~ ".", TRUE ~ E_GT)) %>%
  # replace any B_depth of 0 or B_depth with no EH call with NA for accurate calculation of mean depth
  mutate(B_depth = case_when(B_depth == 0 ~ NA_integer_, E_GT == "./." | E_GT == "." ~ NA_integer_, TRUE ~ B_depth)) %>%
  # separate fields giving base pair/repeat difference from reference into column for each allele
  separate(H_GB,c("H_GB_1_TEMP","H_GB_2_TEMP"), sep="\\|", fill="right", remove=F) %>%
  separate(E_REPCN, c("E_REPCN_1","E_REPCN_2"), sep="/", fill="right", remove=F) %>%
  separate(G_REPCN, c("G_REPCN_1","G_REPCN_2"), sep=",", fill="right", remove=F) %>%
  # split ADSP field to get spanning read depth for Expansion Hunter alleles
  separate(E_ADSP, c("E_ADSP_1","E_ADSP_2"), sep="/", fill="right", remove=F) %>%
  # convert columns from character to numeric where necessary
  mutate_at(c("H_Q","H_DP","H_DFLANKINDEL","H_DSTUTTER", "H_GB_1_TEMP", "H_GB_2_TEMP", "H_GLDIFF",
              "E_REPCN_1","E_REPCN_2","E_ADSP_1","E_ADSP_2",
              "G_Q","G_DP","G_REPCN_1","G_REPCN_2"), as.numeric) %>%
  # add columns to get repeat count difference from reference
  mutate(E_REPDIFF_1 = E_REPCN_1-E_REF, E_REPDIFF_2 = E_REPCN_2-E_REF) %>%
  mutate(G_REPDIFF_1 = G_REPCN_1-G_REF, G_REPDIFF_2 = G_REPCN_2-G_REF) %>%
  # add columns to get fraction of reads that have stutter or flank indels
  mutate(H_STUTTERFRAC = H_DSTUTTER/H_DP, H_FLINDELFRAC = H_DFLANKINDEL/H_DP)

# split MALLREADS and match to each allele from HipSTR call to get spanning read depth (maximum-likelihood-based)
results$H_MALLREADS_1_TEMP <- unlist(as.numeric(mapply(function(x,y){ sub(paste0("^",x,"\\|(.*)"), "\\1", y[grep(paste0("^",x,"\\|"),y)])}, x=results$H_GB_1_TEMP,
y=strsplit(results$H_MALLREADS,";"), SIMPLIFY=FALSE, USE.NAMES=FALSE)))
results$H_MALLREADS_2_TEMP <- unlist(as.numeric(mapply(function(x,y){ sub(paste0("^",x,"\\|(.*)"), "\\1", y[grep(paste0("^",x,"\\|"),y)])}, x=results$H_GB_2_TEMP,
y=strsplit(results$H_MALLREADS,";"), SIMPLIFY=FALSE, USE.NAMES=FALSE)))

# split ALLREADS and match to each allele from HipSTR call to get spanning read depth (alignment-based)
results$H_ALLREADS_1_TEMP <- unlist(as.numeric(mapply(function(x,y){ sub(paste0("^",x,"\\|(.*)"), "\\1", y[grep(paste0("^",x,"\\|"),y)])}, x=results$H_GB_1_TEMP,
y=strsplit(results$H_ALLREADS,";"), SIMPLIFY=FALSE, USE.NAMES=FALSE)))
results$H_ALLREADS_2_TEMP <- unlist(as.numeric(mapply(function(x,y){ sub(paste0("^",x,"\\|(.*)"), "\\1", y[grep(paste0("^",x,"\\|"),y)])}, x=results$H_GB_2_TEMP,
y=strsplit(results$H_ALLREADS,";"), SIMPLIFY=FALSE, USE.NAMES=FALSE)))

# Calculate repeat difference from reference for HipSTR based on sample sex
if (sex == "M") {
  
  results <- results %>%
    # fill in permanent H_MALLREADS columns
    # if on a sex chromosome, fill allele 1 with the allele with the maximum read count
    mutate(H_MALLREADS_1 = 
             if_else((chr == "chrX" | chr == "chrY") & H_MALLREADS_2_TEMP == pmax(H_MALLREADS_1_TEMP, H_MALLREADS_2_TEMP), 
                     H_MALLREADS_2_TEMP, H_MALLREADS_1_TEMP),
           H_MALLREADS_2 =
             if_else((chr == "chrX" | chr == "chrY") & H_MALLREADS_2_TEMP == pmax(H_MALLREADS_1_TEMP, H_MALLREADS_2_TEMP), 
                     H_MALLREADS_1_TEMP, H_MALLREADS_2_TEMP)) %>%
    # fill in permanent H_ALLREADS columns
    # if on a sex chromosome, fill allele 1 with the allele with the maximum read count
    mutate(H_ALLREADS_1 = 
             if_else((chr == "chrX" | chr == "chrY") & H_ALLREADS_2_TEMP == pmax(H_ALLREADS_1_TEMP, H_ALLREADS_2_TEMP), 
                     H_ALLREADS_2_TEMP, H_ALLREADS_1_TEMP),
           H_ALLREADS_2 =
             if_else((chr == "chrX" | chr == "chrY") & H_ALLREADS_2_TEMP == pmax(H_ALLREADS_1_TEMP, H_ALLREADS_2_TEMP), 
                     H_ALLREADS_1_TEMP, H_ALLREADS_2_TEMP)) %>%
    # fill in permanent H_GB columns
    # if on a sex chromosome, fill allele 1 with the allele with the maximum read count
    mutate(H_GB_1 = if_else((chr == "chrX" | chr == "chrY") & H_MALLREADS_2_TEMP == pmax(H_MALLREADS_1_TEMP, H_MALLREADS_2_TEMP), H_GB_2_TEMP, H_GB_1_TEMP),
           H_GB_2 = if_else((chr == "chrX" | chr == "chrY") & H_MALLREADS_2_TEMP == pmax(H_MALLREADS_1_TEMP, H_MALLREADS_2_TEMP), H_GB_1_TEMP, H_GB_2_TEMP)) %>%
    # calculate repeat difference
    # fill in -999 for allele 2 on sex chromosomes to indicate it should not be considered a true call
    mutate(H_REPDIFF_1 = H_GB_1/period.size,
           H_REPDIFF_2 = if_else((chr == "chrX" | chr == "chrY"), -999, H_GB_2/period.size))
  
} else {
  
  # fill in all permanent columns with their corresponding temporary column
  # since sample is female, chrX has 2 alleles and can be filled as autosomes
  # fill in -999 for any chrY calls to indicate it should not be considered a true call
  results <- results %>% 
    mutate(H_MALLREADS_1 = H_MALLREADS_1_TEMP, H_MALLREADS_2 = H_MALLREADS_2_TEMP,
           H_GB_1 = H_GB_1_TEMP, H_GB_2 = H_GB_2_TEMP) %>%
    mutate(H_ALLREADS_1 = H_ALLREADS_1_TEMP, H_ALLREADS_2 = H_ALLREADS_2_TEMP,
           H_GB_1 = H_GB_1_TEMP, H_GB_2 = H_GB_2_TEMP) %>%
    mutate(H_REPDIFF_1 = if_else(chr == "chrY", -999, H_GB_1/period.size), 
           H_REPDIFF_2 = if_else(chr == "chrY", -999, H_GB_2/period.size))

}

# remove temporary columns
results <- results %>% select(-c(H_GB_1_TEMP, H_GB_2_TEMP, H_MALLREADS_1_TEMP, H_MALLREADS_2_TEMP, H_ALLREADS_1_TEMP, H_ALLREADS_2_TEMP))


# split ENCLREADS and match to each allele from GangSTR call to get spanning read depth
results$G_ENCLREADS_1 <- unlist(as.numeric(mapply(function(x,y){ sub(paste0("^",x,",(.*)"), "\\1", y[grep(paste0("^",x,","),y)])}, x=results$G_REPCN_1,
y=strsplit(results$G_ENCLREADS,"\\|"), SIMPLIFY=FALSE, USE.NAMES=FALSE)))
results$G_ENCLREADS_2 <- unlist(as.numeric(mapply(function(x,y){ sub(paste0("^",x,",(.*)"), "\\1", y[grep(paste0("^",x,","),y)])}, x=results$G_REPCN_2,
y=strsplit(results$G_ENCLREADS,"\\|"), SIMPLIFY=FALSE, USE.NAMES=FALSE)))

# Calculate VAF
results <- results %>%
  # Find total spanning read depth in HipSTR by summing read depths in MALLREADS
  mutate(H_MALLREADS_SUM = sapply(strsplit(results$H_MALLREADS,";|\\|"), readdepth)) %>%
  mutate(H_ALLREADS_SUM = sapply(strsplit(results$H_ALLREADS,";|\\|"), readdepth)) %>%
  # Divide by individual read depth to get VAF
  mutate(H_MVAF_1 = H_MALLREADS_1/H_MALLREADS_SUM) %>% mutate(H_MVAF_2 = H_MALLREADS_2/H_MALLREADS_SUM) %>%
  mutate(H_AVAF_1 = H_ALLREADS_1/H_ALLREADS_SUM) %>% mutate(H_AVAF_2 = H_ALLREADS_2/H_ALLREADS_SUM) %>%
  mutate(E_VAF_1 = E_ADSP_1/B_depth) %>% mutate(E_VAF_2 = E_ADSP_2/B_depth) %>%
  # Find total spanning read depth in GangSTR by summing read depths in ENCLREADS
  mutate(G_ENCLREADS_SUM = sapply(strsplit(results$G_ENCLREADS,",|\\|"), readdepth)) %>%
  # Divide by individual read depth to get VAF
  mutate(G_VAF_1 = G_ENCLREADS_1/G_ENCLREADS_SUM) %>% mutate(G_VAF_2 = G_ENCLREADS_2/G_ENCLREADS_SUM)

# Save data frame in RDS file with sample name
saveRDS(results, file=paste0(sampleID,".rds"))
```


### 2. join_str_samples_nextflow.r
```{r}
#!/share/apps/r/4.2.0/gcc/bin/Rscript

#usage: Rscript join_str_samples_nextflow.r ${RDS} ${params.outputbasename}

library(dplyr)
library(tidyr)

args <- commandArgs(trailingOnly = T)

# Initialize results list
results_list <- list()

# Load each RDS file and join data frame to list
for (x in head(args,-1)) {
  results_list[[x]] <- readRDS(x)
}

results <- as_tibble(do.call(rbind, results_list))

# Save list in RDS file
saveRDS(results, file = paste0(tail(args, 1),".rds"))
```




##########################

FILTERING & CONCORDANCE

##########################

# F. GENOTYPE QUALITY FILTERING

## 1. YAML Config
Filtering and concordance calculations will use a YAML config file containing the filtering thresholds and sample information for Mendelian trio concordance calculations.

filter_config.yaml
```{bash}
# quality scores, VAF, and sample fraction are all numbers between 0 and 1
# read counts are any number
HipSTR:
  # all parameters are numbers
  nonA.min.qual: x
  nonA.min.mallreads: x
  nonA.min.allele.mallreads: x
  nonA.min.mvaf: x
  nonA.min.allreads: x
  nonA.min.allele.allreads: x
  nonA.min.avaf: x
  nonA.max.stutter.frac: x
  nonA.max.flankindel.frac: x
  nonA.min.gldiff: x
  nonA.min.mean.mallreads: x
  nonA.min.mean.allreads: x
  nonA.min.mean.qual: x
  nonA.min.sample.frac: x
  A.min.qual: x
  A.min.mallreads: x
  A.min.allele.mallreads: x
  A.min.mvaf: x
  A.min.allreads: x
  A.min.allele.allreads: x
  A.min.avaf: x
  A.max.stutter.frac: x
  A.max.flankindel.frac: x
  A.min.gldiff: x
  A.min.mean.mallreads: x
  A.min.mean.allreads: x
  A.min.mean.qual: x
  A.min.sample.frac: x
GangSTR:
  # all parameters are numbers
  nonA.min.qual: x
  nonA.min.total.reads: x
  nonA.min.allele.reads: x
  nonA.min.vaf: x
  nonA.min.mean.total.reads: x
  nonA.min.mean.qual: x
  nonA.min.sample.frac: x
  A.min.qual: x
  A.min.total.reads: x
  A.min.allele.reads: x
  A.min.vaf: x
  A.min.mean.total.reads: x
  A.min.mean.qual: x
  A.min.sample.frac: x
EH:
  # all parameters are numbers
  nonA.min.total.reads: x
  nonA.min.allele.reads: x
  nonA.min.vaf: x
  nonA.max.vaf: x
  nonA.min.lc: x
  nonA.min.mean.total.reads: x
  nonA.min.sample.frac: x
  A.min.total.reads: x
  A.min.allele.reads: x
  A.min.vaf: x
  A.max.vaf: x
  A.min.lc: x
  A.min.mean.total.reads: x
  A.min.sample.frac: x
ensemble:
  # fill in "HipSTR", "GangSTR", "EH", or "no_caller"
  # caller.one cannot be filled in with "no_caller"
  # caller.two can only be "no_caller" if caller.three is also "no_caller"
  # caller.three can always be "no_caller"
  nonA.caller.one: x
  nonA.caller.two: x
  nonA.caller.three: x
  A.caller.one: x
  A.caller.two: x
  A.caller.three: x
sample:
  # trio ID from Nextflow pipeline
  trio: x
  # sample IDs from Nextflow pipeline
  father: x
  mother: x
  child: x
  # fill in "M" or "F"
  sex.of.child: x

```

## 2. filtering_optimization.r
Code for using YAML config file is commented out in the optimization script. YAML will be used in the final code, but optimization is more efficient using command line arguments.
```{r}
#!/share/apps/r/4.2.0/gcc/bin/Rscript

# usage: filtering_optimization.r [RDS output of join_str_samples_nextflow.r] [condition #] [arguments 3 - 46 (filtering parameters)]
# RDS is output of join_str_samples_nextflow.r with HipSTR, GangSTR, Expansion Hunter, and bedtools data for all loci and all samples in tibble format
# The RDS is the final output of the Nextflow pipeline
# condition number refers to the specific combination of parameters used in that optimization test, for easier tracking of outputs
# arguments 3 - 46 are the filtering parameters used throughout the script

# usage with YAML: [RDS output of join_str_samples_nextflow.r] [YAML config]

# RDS is output of join_str_samples_nextflow.r with HipSTR, GangSTR, Expansion Hunter, and bedtools data for all loci and all samples in tibble format
# The RDS is the final output of the Nextflow pipeline

# YAML CODE IS COMMENTED OUT FOR OPTIMIZATION
# config file is a YAML file with parameter settings for each caller and ensemble calling

library(dplyr)
library(tidyr)
#library(configr)

# load command line arguments
args <- commandArgs(trailingOnly = T)

# Load YAML config file
#config <- read.config(file = args[2])

# load results data frame from RDS
results <- readRDS(args[1])

nonA_loci <- results %>%
  filter(motif.family != "A")

A_loci <- results %>%
  filter(motif.family == "A")


## filtering parameters

# HipSTR

# minimum HipSTR call quality (VCF field Q)
# nonA.min.qual.hipstr <- config$HipSTR$nonA.min.qual
# A.min.qual.hipstr <- config$HipSTR$A.min.qual
nonA.min.qual.hipstr <- args[3]
A.min.qual.hipstr <- args[4]
# minimum total number of reads used to make call in HipSTR (sum of read counts in VCF field MALLREADS)
# nonA.min.mallreads.hipstr <- config$HipSTR$nonA.min.mallreads
# A.min.mallreads.hipstr <- config$HipSTR$A.min.mallreads
nonA.min.mallreads.hipstr <- as.numeric(args[5])
A.min.mallreads.hipstr <- as.numeric(args[6])
# minimum number of reads supporting the called allele in HipSTR (read count for allele matching genotype call in VCF field MALLREADS)
# nonA.min.allele.mallreads.hipstr <- config$HipSTR$nonA.min.allele.mallreads
# A.min.allele.mallreads.hipstr <- config$HipSTR$A.min.allele.mallreads
nonA.min.allele.mallreads.hipstr <- as.numeric(args[7])
A.min.allele.mallreads.hipstr <- as.numeric(args[8])
# minimum variant allele fraction (VAF) for the called allele in HipSTR (MALLREADS for allele / sum of MALLREADS)
# nonA.min.mvaf.hipstr <- config$HipSTR$nonA.min.mvaf
# A.min.mvaf.hipstr <- config$HipSTR$A.min.mvaf
nonA.min.mvaf.hipstr <- as.numeric(args[9])
A.min.mvaf.hipstr <- as.numeric(args[10])
# minimum total number of reads used to make call in HipSTR (sum of read counts in VCF field ALLREADS)
# nonA.min.allreads.hipstr <- config$HipSTR$nonA.min.allreads
# A.min.allreads.hipstr <- config$HipSTR$A.min.allreads
nonA.min.allreads.hipstr <- as.numeric(args[11])
A.min.allreads.hipstr <- as.numeric(args[12])
# minimum number of reads supporting the called allele in HipSTR (read count for allele matching genotype call in VCF field allreads)
# nonA.min.allele.allreads.hipstr <- config$HipSTR$nonA.min.allele.allreads
# A.min.allele.allreads.hipstr <- config$HipSTR$A.min.allele.allreads
nonA.min.allele.allreads.hipstr <- as.numeric(args[13])
A.min.allele.allreads.hipstr <- as.numeric(args[14])
# minimum variant allele fraction (VAF) for the called allele in HipSTR (allreads for allele / sum of ALLREADS)
# nonA.min.avaf.hipstr <- config$HipSTR$nonA.min.avaf
# A.min.avaf.hipstr <- config$HipSTR$A.min.avaf
nonA.min.avaf.hipstr <- as.numeric(args[15])
A.min.avaf.hipstr <- as.numeric(args[16])
# maximum fraction of stutter reads
# nonA.max.stutter.frac.hipstr <- config$HipSTR$nonA.max.stutter.frac
# A.max.stutter.frac.hipstr <- config$HipSTR$A.max.stutter.frac
nonA.max.stutter.frac.hipstr <- as.numeric(args[17])
A.max.stutter.frac.hipstr <- as.numeric(args[18])
# maximum fraction of reads with flank indels
# nonA.max.flankindel.frac.hipstr <- config$HipSTR$nonA.max.flankindel.frac
# A.max.flankindel.frac.hipstr <- config$HipSTR$A.max.flankindel.frac
nonA.max.flankindel.frac.hipstr <- as.numeric(args[19])
A.max.flankindel.frac.hipstr <- as.numeric(args[20])
# minimum difference in likelihood between call and next-best call (GLDIFF)
# nonA.min.gldiff.hipstr <- config$HipSTR$nonA.min.gldiff
# A.min.gldiff.hipstr <- config$HipSTR$A.min.gldiff
nonA.min.gldiff.hipstr <- as.numeric(args[21])
A.min.gldiff.hipstr <- as.numeric(args[22])
# minimum average HipSTR call quality (VCF field Q) across all samples
# nonA.min.mean.qual.hipstr <- config$HipSTR$nonA.min.mean.qual
# A.min.mean.qual.hipstr <- config$HipSTR$A.min.mean.qual
nonA.min.mean.qual.hipstr <- as.numeric(args[23])
A.min.mean.qual.hipstr <- as.numeric(args[24])
# minimum average total number of reads (MALLREADS) used to make call in HipSTR across all samples
# nonA.min.mean.mallreads.hipstr <- config$HipSTR$nonA.min.mean.mallreads
# A.min.mean.mallreads.hipstr <- config$HipSTR$A.min.mean.mallreads
nonA.min.mean.mallreads.hipstr <- as.numeric(args[25])
A.min.mean.mallreads.hipstr <- as.numeric(args[26])
# minimum average total number of reads (ALLREADS) used to make call in HipSTR across all samples
# nonA.min.mean.allreads.hipstr <- config$HipSTR$nonA.min.mean.allreads
# A.min.mean.allreads.hipstr <- config$HipSTR$A.min.mean.allreads
nonA.min.mean.allreads.hipstr <- as.numeric(args[27])
A.min.mean.allreads.hipstr <- as.numeric(args[28])
# minimum fraction of total samples that have a call that passed all sample-level filters
# nonA.min.sample.frac.hipstr <- config$HipSTR$nonA.min.sample.frac
# A.min.sample.frac.hipstr <- config$HipSTR$A.min.sample.frac
nonA.min.sample.frac.hipstr <- as.numeric(args[29])
A.min.sample.frac.hipstr <- as.numeric(args[30])

# GangSTR

# minimum GangSTR call quality (VCF field Q)
# nonA.min.qual.gangstr <- config$GangSTR$nonA.min.qual
# A.min.qual.gangstr <- config$GangSTR$A.min.qual
nonA.min.qual.gangstr <- args[31]
A.min.qual.gangstr <- args[32]
# minimum total number of spanning reads used to make call in GangSTR (sum of read counts in VCF field ENCLREADS)
# nonA.min.total.reads.gangstr <- config$GangSTR$nonA.min.total.reads
# A.min.total.reads.gangstr <- config$GangSTR$A.min.total.reads
nonA.min.total.reads.gangstr <- as.numeric(args[33])
A.min.total.reads.gangstr <- as.numeric(args[34])
# minimum number of reads supporting the called allele in GangSTR (read count for allele matching genotype call in VCF field ENCLREADS)
# nonA.min.allele.reads.gangstr <- config$GangSTR$nonA.min.allele.reads
# A.min.allele.reads.gangstr <- config$GangSTR$A.min.allele.reads
nonA.min.allele.reads.gangstr <- as.numeric(args[35])
A.min.allele.reads.gangstr <- as.numeric(args[36])
# minimum variant allele fraction (VAF) for the called allele in GangSTR (ENCLREADS for allele / sum of ENCLREADS)
# nonA.min.vaf.gangstr <- config$GangSTR$nonA.min.vaf
# A.min.vaf.gangstr <- config$GangSTR$A.min.vaf
nonA.min.vaf.gangstr <- as.numeric(args[37])
A.min.vaf.gangstr <- as.numeric(args[38])
# minimum average GangSTR call quality (VCF field Q) across all samples
# nonA.min.mean.qual.gangstr <- config$GangSTR$nonA.min.mean.qual
# A.min.mean.qual.gangstr <- config$GangSTR$A.min.mean.qual
nonA.min.mean.qual.gangstr <- as.numeric(args[39])
A.min.mean.qual.gangstr <- as.numeric(args[40])
# minimum average total number of reads used to make call in GangSTR across all samples
# nonA.min.mean.total.reads.gangstr <- config$GangSTR$nonA.min.mean.total.reads
# A.min.mean.total.reads.gangstr <- config$GangSTR$A.min.mean.total.reads
nonA.min.mean.total.reads.gangstr <- as.numeric(args[41])
A.min.mean.total.reads.gangstr <- as.numeric(args[42])
# minimum fraction of total samples that have a call that passed all sample-level filters
# nonA.min.sample.frac.gangstr <- config$GangSTR$nonA.min.sample.frac
# A.min.sample.frac.gangstr <- config$GangSTR$A.min.sample.frac
nonA.min.sample.frac.gangstr <- as.numeric(args[43])
A.min.sample.frac.gangstr <- as.numeric(args[44])

# Expansion Hunter

# minimum number of spanning reads counted by bedtools coverage at each locus
# nonA.min.total.reads.eh <- config$EH$nonA.min.total.reads
# A.min.total.reads.eh <- config$EH$A.min.total.reads
nonA.min.total.reads.eh <- as.numeric(args[45])
A.min.total.reads.eh <- as.numeric(args[46])
# minimum number of reads supporting the called allele in EH (read count for allele matching genotype call in VCF field ADSP)
# nonA.min.allele.reads.eh <- config$EH$nonA.min.allele.reads
# A.min.allele.reads.eh <- config$EH$A.min.allele.reads
nonA.min.allele.reads.eh <- as.numeric(args[47])
A.min.allele.reads.eh <- as.numeric(args[48])
# minimum variant allele fraction (VAF) for the called allele in EH (ADSP for allele / bedtools coverage spanning reads)
# nonA.min.vaf.eh <- config$EH$nonA.min.vaf
# A.min.vaf.eh <- config$EH$A.min.vaf
nonA.min.vaf.eh <- as.numeric(args[49])
A.min.vaf.eh <- as.numeric(args[50])
# maximum variant allele fraction (VAF) for the called allele in EH (ADSP for allele / bedtools coverage spanning reads)
# nonA.max.vaf.eh <- config$EH$nonA.max.vaf
# A.max.vaf.eh <- config$EH$A.max.vaf
nonA.max.vaf.eh <- as.numeric(args[51])
A.max.vaf.eh <- as.numeric(args[52])
# minimum average locus coverage (LC) in EH
# nonA.min.lc.eh <- config$EH$nonA.min.lc
# A.min.lc.eh <- config$EH$A.min.lc
nonA.min.lc.eh <- as.numeric(args[53])
A.min.lc.eh <- as.numeric(args[54])
# minimum average number of spanning reads counted by bedtools coverage at each locus across all samples
# nonA.min.mean.total.reads.eh <- config$EH$nonA.min.mean.total.reads
# A.min.mean.total.reads.eh <- config$EH$A.min.mean.total.reads
nonA.min.mean.total.reads.eh <- as.numeric(args[55])
A.min.mean.total.reads.eh <- as.numeric(args[56])
# minimum fraction of total samples that have a call that passed all sample-level filters
# nonA.min.sample.frac.eh <- config$EH$nonA.min.sample.frac
# A.min.sample.frac.eh <- config$EH$A.min.sample.frac
nonA.min.sample.frac.eh <- as.numeric(args[57])
A.min.sample.frac.eh <- as.numeric(args[58])

# ensemble

# 1st preferred caller (HipSTR, GangSTR, or EH)
# nonA.caller.one <- config$ensemble$nonA.caller.one
# A.caller.one <- config$ensemble$A.caller.one
nonA.caller.one <- args[59]
A.caller.one <- args[60]
# 2nd preferred caller (HipSTR, GangSTR, or EH)
# nonA.caller.two <- config$ensemble$nonA.caller.two
# A.caller.two <- config$ensemble$A.caller.two
nonA.caller.two <- args[61]
A.caller.two <- args[62]
# 3rd preferred caller (HipSTR, GangSTR, or EH)
# nonA.caller.three <- config$ensemble$nonA.caller.three
# A.caller.three <- config$ensemble$A.caller.three
nonA.caller.three <- args[63]
A.caller.three <- args[64]



filtering_function <-
  function(subset_data,
           min.qual.hipstr, min.mallreads.hipstr, min.allreads.hipstr,
           min.allele.mallreads.hipstr, min.allele.allreads.hipstr, min.mvaf.hipstr, min.avaf.hipstr,
           max.stutter.frac.hipstr, max.flankindel.frac.hipstr, min.gldiff.hipstr,
           min.mean.qual.hipstr, min.mean.mallreads.hipstr, min.mean.allreads.hipstr, min.sample.frac.hipstr,
           min.qual.gangstr, min.total.reads.gangstr, min.allele.reads.gangstr, min.vaf.gangstr,
           min.mean.qual.gangstr, min.mean.total.reads.gangstr, min.sample.frac.gangstr,
           min.total.reads.eh, min.allele.reads.eh, min.vaf.eh, max.vaf.eh, min.lc.eh,
           min.mean.total.reads.eh, min.sample.frac.eh,
           caller.one, caller.two, caller.three) {

  ## Sample-level filters

  # add column for each caller
  # fill with TRUE if sample passes all filters for that caller and chromosome
  # fill with FALSE if sample fails any filter for that caller

  if((caller.one == "no_caller")) {
    stop("caller.one cannot be set to 'no_caller'")
  }
  
  if((caller.two == "no_caller" & caller.three != "no_caller")){
    stop("If caller.two == 'no_caller', caller.three must also be set to 'no_caller'")
  }
    
  subset_data <- subset_data %>%
  # HipSTR
  mutate(HipSTR_passsample =
           # if on chrX or chrY and male
           case_when(sex == "M" & (chr == "chrX" | chr == "chrY") &
                     # if there is a genotype call
                     H_GT != "." & !is.na(H_GT) &
                     # quality
                     H_Q >= min.qual.hipstr & !is.na(H_Q) &
                     # total read depth
                     H_MALLREADS_SUM >= min.mallreads.hipstr & !is.na(H_MALLREADS_SUM) &
                     H_ALLREADS_SUM >= min.allreads.hipstr & !is.na(H_ALLREADS_SUM) &                       
                     # allele 1 read depth
                     H_MALLREADS_1 >= min.allele.mallreads.hipstr & !is.na(H_MALLREADS_1) &
                     H_ALLREADS_1 >= min.allele.allreads.hipstr & !is.na(H_ALLREADS_1) &                       
                     # allele 1 VAF
                     H_MVAF_1 >= min.mvaf.hipstr & !is.na(H_MVAF_1) &
                     H_AVAF_1 >= min.avaf.hipstr & !is.na(H_AVAF_1) &                       
                     # stutter fraction
                     H_STUTTERFRAC <= max.stutter.frac.hipstr & !is.na(H_STUTTERFRAC) &
                     # flank indel fraction
                     H_FLINDELFRAC <= max.flankindel.frac.hipstr & !is.na(H_FLINDELFRAC) &
                     # likelihood difference
                     # GLDIFF is sometimes not reported for loci with genotypes
                     # We do not want to throw out those loci                       
                     (H_GLDIFF >= min.gldiff.hipstr | is.na(H_GLDIFF)) ~ TRUE,

                     # if on an autosome and male or on any chromosome other than chrY and female
                     ((sex == "M" & chr != "chrX" & chr != "chrY") | (sex == "F" & chr != "chrY")) &
                     # if there is a genotype call
                     H_GT != "." & !is.na(H_GT) &
                     # quality
                     H_Q >= min.qual.hipstr & !is.na(H_Q) &
                     # total read depth
                     H_MALLREADS_SUM >= min.mallreads.hipstr & !is.na(H_MALLREADS_SUM) &
                     H_ALLREADS_SUM >= min.allreads.hipstr & !is.na(H_ALLREADS_SUM) &                       
                     # allele 1 read depth
                     H_MALLREADS_1 >= min.allele.mallreads.hipstr & !is.na(H_MALLREADS_1) &
                     H_ALLREADS_1 >= min.allele.allreads.hipstr & !is.na(H_ALLREADS_1) &                       
                     # allele 2 read depth
                     H_MALLREADS_2 >= min.allele.mallreads.hipstr & !is.na(H_MALLREADS_2) &
                     H_ALLREADS_2 >= min.allele.allreads.hipstr & !is.na(H_ALLREADS_2) &                       
                     # allele 1 VAF
                     H_MVAF_1 >= min.mvaf.hipstr & !is.na(H_MVAF_1) &
                     H_AVAF_1 >= min.avaf.hipstr & !is.na(H_AVAF_1) &                       
                     # allele 2 VAF
                     H_MVAF_2 >= min.mvaf.hipstr & !is.na(H_MVAF_2) &
                     H_AVAF_2 >= min.avaf.hipstr & !is.na(H_AVAF_2) &                       
                     # stutter fraction
                     H_STUTTERFRAC <= max.stutter.frac.hipstr & !is.na(H_STUTTERFRAC) &
                     # flank indel fraction
                     H_FLINDELFRAC <= max.flankindel.frac.hipstr & !is.na(H_FLINDELFRAC) &
                     # likelihood difference
                     # GLDIFF is sometimes not reported for loci with genotypes
                     # We do not want to throw out those loci
                     (H_GLDIFF >= min.gldiff.hipstr | is.na(H_GLDIFF)) ~ TRUE,

                     # if any filter failed
                     TRUE ~ FALSE)) %>%
    
  # GangSTR
  mutate(GangSTR_passsample =
           # if on chrX or chr Y and male
           case_when(sex == "M" & (chr == "chrX" | chr == "chrY") &
                     # if there is a genotype call
                     G_GT != "." & !is.na(G_GT) &
                     # quality
                     G_Q >= min.qual.gangstr & !is.na(G_Q) &
                     # total read depth
                     G_ENCLREADS_SUM >= min.total.reads.gangstr & !is.na(G_ENCLREADS_SUM) &
                     # allele 1 read depth
                     G_ENCLREADS_1 >= min.allele.reads.gangstr & !is.na(G_ENCLREADS_1) &
                     # allele 1 VAF
                     G_VAF_1 >= min.vaf.gangstr & !is.na(G_VAF_1) ~ TRUE,

                     # if on an autosome and male or on any chromosome other than chrY and female
                     ((sex == "M" & chr != "chrX" & chr != "chrY") | (sex == "F" & chr != "chrY")) &
                     # if there is a genotype call
                     G_GT != "." & !is.na(G_GT) &
                     # quality
                     G_Q >= min.qual.gangstr & !is.na(G_Q) &
                     # total read depth
                     G_ENCLREADS_SUM >= min.total.reads.gangstr & !is.na(G_ENCLREADS_SUM) &
                     # allele 1 read depth
                     G_ENCLREADS_1 >= min.allele.reads.gangstr & !is.na(G_ENCLREADS_1) &
                     # allele 2 read depth
                     G_ENCLREADS_2 >= min.allele.reads.gangstr & !is.na(G_ENCLREADS_2) &
                     # allele 1 VAF
                     G_VAF_1 >= min.vaf.gangstr & !is.na(G_VAF_1) &
                     # allele 2 VAF
                     G_VAF_2 >= min.vaf.gangstr & !is.na(G_VAF_2)~ TRUE,

                     # if any filter failed
                     TRUE ~ FALSE)) %>%
    
  # Expansion Hunter
  mutate(EH_passsample =
           # if on chrX or chrY and male
           case_when(sex == "M" & (chr == "chrX" | chr == "chrY") &
                    # if there is a genotype call
                    E_GT != "." & !is.na(E_GT) &
                    # alleles supported by spanning reads only
                    (E_SO == "SPANNING" | E_SO == "SPANNING/SPANNING") & !is.na(E_SO) &
                    # total read depth (bedtools)
                    B_depth >= min.total.reads.eh & !is.na(B_depth) &
                    # allele 1 read depth
                    E_ADSP_1 >= min.allele.reads.eh & !is.na(E_ADSP_1) &
                    # allele 1 VAF
                    E_VAF_1 >= min.vaf.eh & E_VAF_1 <= max.vaf.eh & !is.na(E_VAF_1) &
                    # average locus coverage
                    E_LC >= min.lc.eh & !is.na(E_LC) ~ TRUE,

                    # if on an autosome and male or on any chromosome other than chrY and female
                    ((sex == "M" & chr != "chrX" & chr != "chrY") | (sex == "F" & chr != "chrY")) &
                    # if there is a genotype call
                    E_GT != "./." & !is.na(E_GT) &
                    # alleles supported by spanning reads only
                    (E_SO == "SPANNING" | E_SO == "SPANNING/SPANNING") & !is.na(E_SO) &
                    # total read depth (bedtools)
                    B_depth >= min.total.reads.eh & !is.na(B_depth) &
                    # allele 1 read depth
                    E_ADSP_1 >= min.allele.reads.eh & !is.na(E_ADSP_1) &
                    # allele 2 read depth
                    E_ADSP_2 >= min.allele.reads.eh & !is.na(E_ADSP_2) &
                    # allele 1 VAF
                    E_VAF_1 >= min.vaf.eh & E_VAF_1 <= max.vaf.eh & !is.na(E_VAF_1) &
                    # allele 2 VAF
                    E_VAF_2 >= min.vaf.eh & E_VAF_2 <= max.vaf.eh & !is.na(E_VAF_2) &
                    # average locus coverage
                    E_LC >= min.lc.eh & !is.na(E_LC) ~ TRUE,

                    # if any filter failed
                    TRUE ~ FALSE))

  ## Locus-level filters

  # get number of samples from table
  num_samples = n_distinct(subset_data$sampleID)

  # Calculate and apply locus-level filters
  depth_qual_avg <- subset_data %>% group_by(name) %>%
    summarise(# average quality
              mean_qual_H = mean(H_Q, na.rm = T),
              mean_qual_G = mean(G_Q, na.rm = T),
              # average total read depth
              mean_mdepth_H = mean(H_MALLREADS_SUM, na.rm = T),
              mean_adepth_H = mean(H_ALLREADS_SUM, na.rm = T),
              mean_depth_G = mean(G_ENCLREADS_SUM, na.rm = T),
              mean_depth_E = mean(B_depth, na.rm = T),
              #HipSTR filter
              locus_means_H = if_else(mean_qual_H >= min.mean.qual.hipstr & 
                                        mean_mdepth_H >= min.mean.mallreads.hipstr &
                                        mean_adepth_H >= min.mean.allreads.hipstr, T, F),
              # GangSTR filter
              locus_means_G = if_else(mean_qual_G >= min.mean.qual.gangstr & mean_depth_G >= min.mean.total.reads.gangstr, T, F),
              # Expansion Hunter filter
              locus_means_E = if_else(mean_depth_E >= min.mean.total.reads.eh, T, F)) %>%
    # replace NA values in locus means columns with FALSE
    mutate(locus_means_H = if_else(is.na(locus_means_H) == T, F, locus_means_H),
           locus_means_G = if_else(is.na(locus_means_G) == T, F, locus_means_G),
           locus_means_E = if_else(is.na(locus_means_E) == T, F, locus_means_E))
  

  # Calculate the fraction of samples in the data frame that passed the sample-level filters for each locus and caller
  # If the sample passed sample-level filters, the column [HipSTR|GangSTR|EH]_passsample in the results data frame == TRUE
  # The number of samples passing sample-level filters at each locus is found using summarise(sum([HipSTR|GangSTR|EH]_passsample == TRUE))
  # Then that number is divided by the total number of samples to get the fraction of samples passing at each locus
  # If the resulting fraction is > min.sample.frac.[hipstr|gangstr|eh], sample_frac_filter_[H|G|E] == TRUE for that locus
  # If the resulting fraction is < min.sample.frac.[hipstr|gangstr|eh], sample_frac_filter_[H|G|E] == FALSE for that locus
  sample_frac_filter <- subset_data %>% select(c(name, sampleID, HipSTR_passsample, GangSTR_passsample, EH_passsample)) %>%
    group_by(name) %>%
    # HipSTR
    summarise(sample_frac_filter_H = if_else(sum(HipSTR_passsample)/num_samples >= min.sample.frac.hipstr, T, F),
              # GangSTR
              sample_frac_filter_G = if_else(sum(GangSTR_passsample)/num_samples >= min.sample.frac.gangstr, T, F),
              # EH
              sample_frac_filter_E = if_else(sum(EH_passsample)/num_samples >= min.sample.frac.eh, T, F))


  # Join to main table
  subset_data <- left_join(subset_data, sample_frac_filter, by = "name") %>% left_join(., depth_qual_avg, by = "name")

  # Use Boolean logic to decide whether to keep locus
  subset_data <- subset_data %>%
    mutate(HipSTR_passlocus = locus_means_H & sample_frac_filter_H,
          GangSTR_passlocus = locus_means_G & sample_frac_filter_G,
          EH_passlocus = locus_means_E & sample_frac_filter_E)
  
  # Label caller.one calls, and set remaining to "no call"
  subset_data <- subset_data %>% mutate(caller = if_else(!!as.name(paste0(caller.one,"_passlocus")) == T,caller.one,"no call"))

  # Label caller.two calls
  if(caller.two != "no_caller"){
    subset_data <- subset_data %>% mutate(caller=if_else(caller == "no call" &    !!as.name(paste0(caller.two,"_passlocus")) == T,caller.two,caller))
  }
  
  # Label caller.three calls
  if(caller.three != "no_caller"){
    subset_data <- subset_data %>% mutate(caller=if_else(caller == "no call" & !!as.name(paste0(caller.three,"_passlocus")) == T,caller.three,caller))
  }


  # Extract genotypes for all loci with a passing call
  
  # Fill in allele_1
  # Genotypes for caller.one calls, and set remaining to NA
  subset_data <- subset_data %>% mutate(allele_1 = if_else(caller == caller.one & !!as.name(paste0(caller.one,"_passsample")) == T, !!as.name(paste0(substr(caller.one,1,1),"_REPDIFF_1")), NA_real_))
  # Genotypes for caller.two calls
  if(caller.two != "no_caller") {
    subset_data <- subset_data %>% mutate(allele_1 = if_else(caller == caller.two & !!as.name(paste0(caller.two,"_passsample")) == T, !!as.name(paste0(substr(caller.two,1,1),"_REPDIFF_1")), allele_1))
  }
  # Genotypes for caller.three calls
  if(caller.three != "no_caller") {
    subset_data <- subset_data %>% mutate(allele_1 = if_else(caller == caller.three & !!as.name(paste0(caller.three,"_passsample")) == T, !!as.name(paste0(substr(caller.three,1,1),"_REPDIFF_1")), allele_1))
  }
  
  # Fill in allele_2
  # Genotypes for caller.one calls, and set remaining to NA
  subset_data <- subset_data %>% mutate(allele_2 = if_else(caller == caller.one & !!as.name(paste0(caller.one,"_passsample")) == T, !!as.name(paste0(substr(caller.one,1,1),"_REPDIFF_2")), NA_real_))
  # Genotypes for caller.two calls
  if(caller.two != "no_caller") {
    subset_data <- subset_data %>% mutate(allele_2 = if_else(caller == caller.two & !!as.name(paste0(caller.two,"_passsample")) == T, !!as.name(paste0(substr(caller.two,1,1),"_REPDIFF_2")), allele_2))
  }
  # Genotypes for caller.three calls
  if(caller.three != "no_caller") {
    subset_data <- subset_data %>% mutate(allele_2 = if_else(caller == caller.three & !!as.name(paste0(caller.three,"_passsample")) == T, !!as.name(paste0(substr(caller.three,1,1),"_REPDIFF_2")), allele_2))
  }
  
  
  return(subset_data)

} # END OF FILTERING FUNCTION

# Run filtering on non-poly-A loci
nonA_loci <-
  filtering_function(nonA_loci,
           nonA.min.qual.hipstr, nonA.min.mallreads.hipstr, nonA.min.allreads.hipstr,
           nonA.min.allele.mallreads.hipstr, nonA.min.allele.allreads.hipstr, nonA.min.mvaf.hipstr, nonA.min.avaf.hipstr,
           nonA.max.stutter.frac.hipstr, nonA.max.flankindel.frac.hipstr, nonA.min.gldiff.hipstr,
           nonA.min.mean.qual.hipstr, nonA.min.mean.mallreads.hipstr, nonA.min.mean.allreads.hipstr, nonA.min.sample.frac.hipstr,
           nonA.min.qual.gangstr, nonA.min.total.reads.gangstr, nonA.min.allele.reads.gangstr, nonA.min.vaf.gangstr,
           nonA.min.mean.qual.gangstr, nonA.min.mean.total.reads.gangstr, nonA.min.sample.frac.gangstr,
           nonA.min.total.reads.eh, nonA.min.allele.reads.eh, nonA.min.vaf.eh, nonA.max.vaf.eh, nonA.min.lc.eh,
           nonA.min.mean.total.reads.eh, nonA.min.sample.frac.eh,
           nonA.caller.one, nonA.caller.two, nonA.caller.three)


# Run filtering on poly-A loci
A_loci <-
  filtering_function(A_loci,
           A.min.qual.hipstr, A.min.mallreads.hipstr, A.min.allreads.hipstr,
           A.min.allele.mallreads.hipstr, A.min.allele.allreads.hipstr, A.min.mvaf.hipstr, A.min.avaf.hipstr,
           A.max.stutter.frac.hipstr, A.max.flankindel.frac.hipstr, A.min.gldiff.hipstr,
           A.min.mean.qual.hipstr, A.min.mean.mallreads.hipstr, A.min.mean.allreads.hipstr, A.min.sample.frac.hipstr,
           A.min.qual.gangstr, A.min.total.reads.gangstr, A.min.allele.reads.gangstr, A.min.vaf.gangstr,
           A.min.mean.qual.gangstr, A.min.mean.total.reads.gangstr, A.min.sample.frac.gangstr,
           A.min.total.reads.eh, A.min.allele.reads.eh, A.min.vaf.eh, A.max.vaf.eh, A.min.lc.eh,
           A.min.mean.total.reads.eh, A.min.sample.frac.eh,
           A.caller.one, A.caller.two, A.caller.three)

# Define chromosome sort order
chrOrder <-c(paste0("chr",1:22),"chrX", "chrY")

# Join non-A and poly-A loci and sort by chromosome and start coordinate
filtered_results <- rbind(nonA_loci, A_loci) %>%
  arrange(sampleID, factor(chr, chrOrder), start)

# save genotypes list in RDS
#saveRDS(filtered_results, file = paste0(sub('\\.rds$', '', args[1]),"_genotypes.rds"))
saveRDS(filtered_results, file = paste0(sub('\\.rds$', '', args[1]),"_",args[2],"_genotypes.rds"))

# save genotypes and metadata in a CSV
#write.csv(filtered_results, file = paste0(sub('\\.rds$', '', args[1]),"_genotypes.csv") , row.names = F, quote = F)
write.csv(filtered_results, file = paste0(sub('\\.rds$', '', args[1]),"_",args[2],"_genotypes.csv") , row.names = F, quote = F)

```


# H. MENDELIAN CONCORDANCE CALCULATION (1 trio)

## 1. trio_concordance.r
```{r}
#!/share/apps/r/4.2.0/gcc/bin/Rscript

# usage: Rscript trio_concordance.r [genotypes RDS] [config file]

# genotypes RDS is output of output_genotypes.r with genotype data for trios
# config file is a YAML file with parameter settings for each caller and ensemble calling

# CONCORDANCE LOGIC:
# If the locus has genotypes for all samples and is concordant, concordance == TRUE
# If the locus has genotypes for all samples and is discordant, concordance == FALSE
# If the locus is missing a genotype for any sample, concordance == "missing genotype"
## Father can be missing genotype for chrX concordance in trios with male child without disrupting concordance calculation
## Mother can be missing genotype for chrY concordance in trios with male child without disrupting concordance calculation
# When concordance column is joined back to full list of loci, any locus that did not pass filtering will have concordance == NA

library(dplyr)
library(tidyr)
library(configr)

args <- commandArgs(trailingOnly = T)

# Load genotypes list from RDS
results <- readRDS(args[1])

# Load YAML config file
config <- read.config(file = args[2])

# filter genotypes for trio of interest
results <- results %>% filter(trio == config$sample$trio)


# save sample name as variable
father_1 <- paste0("allele_1_",config$sample$father)
father_2 <- paste0("allele_2_",config$sample$father)
mother_1 <- paste0("allele_1_",config$sample$mother)
mother_2 <- paste0("allele_2_",config$sample$mother)
child_1 <- paste0("allele_1_",config$sample$child)
child_2 <- paste0("allele_2_",config$sample$child)


# Prepare results data frame for concordance calculation
trio <- results %>%
  # filter for loci with calls
  filter(caller %in% c("HipSTR","GangSTR","EH")) %>%
  # keep call data and some metadata (for joining)
  select(c(name, chr, sampleID, allele_1, allele_2)) %>%
  pivot_wider(names_from = sampleID, values_from = c(allele_1, allele_2))


# Determine concordance at each locus
# Concordance logic is dependent on the sex of the child in the trio
if (config$sample$sex.of.child == "M") {

  # calculate by row
  trio <- trio %>% rowwise() %>%
    mutate(concordance =
           case_when(
           # if the chromosome is X and call is concordant
             chr == "chrX" &
             # if no genotypes are missing
             !anyNA(c(get(mother_1),get(mother_2),get(child_1))) &
             # is either maternal allele present in the child?
             (get(mother_1) == get(child_1) | get(mother_2) == get(child_1)) &
             # is the first child allele present in the mother?
             # second allele on X chromosome is always -999 in males
             get(child_1) %in% c(get(mother_1),get(mother_2)) ~ TRUE,
             
           # if the chromosome is Y and call is concordant
             chr == "chrY" &
             # if no genotypes are missing
             !anyNA(c(get(father_1),get(child_1))) &
             # is the first paternal allele present in the child?
             get(father_1) == get(child_1) ~ TRUE,

           # if the chromosome is an autosome and call is concordant
             chr != "chrX" & chr != "chrY" &
             # if no genotypes are missing
             !anyNA(c(get(father_1),get(father_2),get(mother_1),get(mother_2),get(child_1),get(child_2))) &
             # is either paternal allele present in the child?
             (get(father_1) %in% c(get(child_1),get(child_2)) | get(father_2) %in% c(get(child_1),get(child_2))) &
             # is either maternal allele present in the child?
             (get(mother_1) %in% c(get(child_1),get(child_2)) | get(mother_2) %in% c(get(child_1),get(child_2))) &
             # is the first child allele present in at least one parent?
             get(child_1) %in% c(get(father_1),get(father_2),get(mother_1),get(mother_2)) &
             # is the second child allele present in at least one parent?
             get(child_2) %in% c(get(father_1),get(father_2),get(mother_1),get(mother_2)) ~ TRUE,

           
           # if the chromosome is X and any maternal allele or child allele is missing
             chr == "chrX" & anyNA(c(get(mother_1),get(mother_2),get(child_1))) ~ NA,
           
           # if the chromosome is Y and the paternal or child allele is missing
             chr == "chrY" & anyNA(c(get(father_1),get(child_1))) ~ NA,
           
           # if the chromosome is an autosome and any allele is missing
             chr != "chrX" & chr != "chrY" & anyNA(c(get(father_1),get(father_2),get(mother_1),get(mother_2),get(child_1),get(child_2))) ~ NA,

           # if the call is discordant
             TRUE ~ FALSE))

} else {

  # calculate by row
  trio <- trio %>% rowwise() %>%
    mutate(concordance =
           case_when(
           # if the chromosome is X and the call is concordant
             chr == "chrX" &
             # if no genotypes are missing
             !anyNA(c(get(father_1),get(mother_1),get(mother_2),get(child_1),get(child_2))) &
             # is either paternal allele present in the child?
             get(father_1) %in% c(get(child_1),get(child_2)) &
             # is either maternal allele present in the child?
             (get(mother_1) %in% c(get(child_1),get(child_2)) | get(mother_2) %in% c(get(child_1),get(child_2))) &
             # is the first child allele present in at least one parent?
             get(child_1) %in% c(get(father_1),get(mother_1),get(mother_2)) &
             # is the second child allele present in at least one parent?
             get(child_2) %in% c(get(father_1),get(mother_1),get(mother_2)) ~ TRUE,
             
           # if the chromosome is an autosome and the call is concordant
             chr != "chrX" & chr != "chrY" &
             # if no genotypes are missing
             !anyNA(c(get(father_1),get(father_2),get(mother_1),get(mother_2),get(child_1),get(child_2))) &
             # is either paternal allele present in the child?
             (get(father_1) %in% c(get(child_1),get(child_2)) | get(father_2) %in% c(get(child_1),get(child_2))) &
             # is either maternal allele present in the child?
             (get(mother_1) %in% c(get(child_1),get(child_2)) | get(mother_2) %in% c(get(child_1),get(child_2))) &
             # is the first child allele present in at least one parent?
             get(child_1) %in% c(get(father_1),get(father_2),get(mother_1),get(mother_2)) &
             # is the second child allele present in at least one parent?
             get(child_2) %in% c(get(father_1),get(father_2),get(mother_1),get(mother_2)) ~ TRUE,
           
           # if the chromosome is Y, disregard call because chrY concordance isn't possible with a female child
             chr == "chrY" ~ NA,

           # if the chromosome is X and either maternal allele, either child allele, or the paternal allele are missing
             chr == "chrX" & anyNA(c(get(father_1),get(mother_1),get(mother_2),get(child_1),get(child_2))) ~ NA,
           
           # if the chromosome is an autosome and any allele is missing
             chr != "chrX" & chr != "chrY" & anyNA(c(get(father_1),get(father_2),get(mother_1),get(mother_2),get(child_1),get(child_2))) ~ NA,

           # if the call is discordant,
             TRUE ~ FALSE))
}


results <- left_join(results, select(trio,c(name,concordance)), by = c("name"))

nonA_results_stats <- results %>% group_by(name, chr, period.size, motif.family, caller, concordance) %>% summarise() %>% filter(motif.family != "A")
A_results_stats <- results %>% group_by(name, chr, period.size, motif.family, caller, concordance) %>% summarise() %>% filter(motif.family == "A")

# Calculate statistics for concordance and filtering data

# non poly-A loci stats
nonA_stats <-
  data.frame(
    # DELETE condition column when finished with optimization since parameters will not change
    condition = sub('.*_([0-9]+)_genotypes\\.rds$', "\\1",args[1]),
    nonA_num_loci_panel = nrow(nonA_results_stats),
    nonA_num_2bp_loci = length(which(nchar(nonA_results_stats$motif.family) == 2)),
    nonA_num_3bp_loci = length(which(nchar(nonA_results_stats$motif.family) == 3)),
    nonA_num_4bp_loci = length(which(nchar(nonA_results_stats$motif.family) == 4)),
    nonA_num_loci_passing_filters = length(which(nonA_results_stats$caller != "no call")),
    nonA_num_2bp_loci_passing_filters = length(which(nchar(nonA_results_stats$motif.family) == 2 & nonA_results_stats$caller != "no call")),
    nonA_num_3bp_loci_passing_filters = length(which(nchar(nonA_results_stats$motif.family) == 3 & nonA_results_stats$caller != "no call")),
    nonA_num_4bp_loci_passing_filters = length(which(nchar(nonA_results_stats$motif.family) == 4 & nonA_results_stats$caller != "no call")),
    nonA_num_loci_failing_filters = length(which(nonA_results_stats$caller == "no call")),
    nonA_num_2bp_loci_failing_filters = length(which(nchar(nonA_results_stats$motif.family) == 2 & nonA_results_stats$caller == "no call")),
    nonA_num_3bp_loci_failing_filters = length(which(nchar(nonA_results_stats$motif.family) == 3 & nonA_results_stats$caller == "no call")),
    nonA_num_4bp_loci_failing_filters = length(which(nchar(nonA_results_stats$motif.family) == 4 & nonA_results_stats$caller == "no call")),
    nonA_num_loci_caller_hipstr = length(which(nonA_results_stats$caller == "HipSTR")),
    nonA_num_loci_caller_gangstr = length(which(nonA_results_stats$caller == "GangSTR")),
    nonA_num_loci_caller_eh = length(which(nonA_results_stats$caller == "EH")),
    nonA_num_loci_fully_genotyped = length(which(nonA_results_stats$concordance == TRUE | nonA_results_stats$concordance == FALSE)),
    nonA_num_2bp_loci_fully_genotyped = length(which((nonA_results_stats$concordance == TRUE | nonA_results_stats$concordance == FALSE)
                                                      & nchar(nonA_results_stats$motif.family) == 2)),
    nonA_num_3bp_loci_fully_genotyped = length(which((nonA_results_stats$concordance == TRUE | nonA_results_stats$concordance == FALSE)
                                                      & nchar(nonA_results_stats$motif.family) == 3)),
    nonA_num_4bp_loci_fully_genotyped = length(which((nonA_results_stats$concordance == TRUE | nonA_results_stats$concordance == FALSE)
                                                      & nchar(nonA_results_stats$motif.family) == 4)),
    nonA_num_hipstr_loci_fully_genotyped = length(which((nonA_results_stats$concordance == TRUE | nonA_results_stats$concordance == FALSE)
                                                          & nonA_results_stats$caller == "HipSTR")),
    nonA_num_gangstr_loci_fully_genotyped = length(which((nonA_results_stats$concordance == TRUE | nonA_results_stats$concordance == FALSE)
                                                          & nonA_results_stats$caller == "GangSTR")),
    nonA_num_eh_loci_fully_genotyped = length(which((nonA_results_stats$concordance == TRUE | nonA_results_stats$concordance == FALSE)
                                                      & nonA_results_stats$caller == "EH")),
    nonA_num_concordant_loci = length(which(nonA_results_stats$concordance == TRUE)),
    nonA_num_2bp_concordant_loci = length(which(nchar(nonA_results_stats$motif.family) == 2 & nonA_results_stats$concordance == TRUE)),
    nonA_num_3bp_concordant_loci = length(which(nchar(nonA_results_stats$motif.family) == 3 & nonA_results_stats$concordance == TRUE)),
    nonA_num_4bp_concordant_loci = length(which(nchar(nonA_results_stats$motif.family) == 4 & nonA_results_stats$concordance == TRUE)),
    nonA_num_hipstr_concordant_loci = length(which(nonA_results_stats$caller == "HipSTR" & nonA_results_stats$concordance == TRUE)),
    nonA_num_gangstr_concordant_loci = length(which(nonA_results_stats$caller == "GangSTR" & nonA_results_stats$concordance == TRUE)),
    nonA_num_eh_concordant_loci = length(which(nonA_results_stats$caller == "EH" & nonA_results_stats$concordance == TRUE)),
    nonA_num_discordant_loci = length(which(nonA_results_stats$concordance == FALSE)),
    nonA_num_2bp_discordant_loci = length(which(nchar(nonA_results_stats$motif.family) == 2 & nonA_results_stats$concordance == FALSE)),
    nonA_num_3bp_discordant_loci = length(which(nchar(nonA_results_stats$motif.family) == 3 & nonA_results_stats$concordance == FALSE)),
    nonA_num_4bp_discordant_loci = length(which(nchar(nonA_results_stats$motif.family) == 4 & nonA_results_stats$concordance == FALSE)),
    nonA_num_hipstr_discordant_loci = length(which(nonA_results_stats$caller == "HipSTR" & nonA_results_stats$concordance == FALSE)),
    nonA_num_gangstr_discordant_loci = length(which(nonA_results_stats$caller == "GangSTR" & nonA_results_stats$concordance == FALSE)),
    nonA_num_eh_discordant_loci = length(which(nonA_results_stats$caller == "EH" & nonA_results_stats$concordance == FALSE)),
    nonA_num_loci_passing_filter_missing_gt = length(which(is.na(nonA_results_stats$concordance) & nonA_results_stats$caller != "no call")),
    nonA_num_2bp_loci_passing_filter_missing_gt = length(which(nchar(nonA_results_stats$motif.family) == 2 & is.na(nonA_results_stats$concordance) & nonA_results_stats$caller != "no call")),
    nonA_num_3bp_loci_passing_filter_missing_gt = length(which(nchar(nonA_results_stats$motif.family) == 3 & is.na(nonA_results_stats$concordance) & nonA_results_stats$caller != "no call")),
    nonA_num_4bp_loci_passing_filter_missing_gt = length(which(nchar(nonA_results_stats$motif.family) == 4 & is.na(nonA_results_stats$concordance & nonA_results_stats$caller != "no call"))),
    nonA_num_hipstr_loci_passing_filter_missing_gt = length(which(nonA_results_stats$caller == "HipSTR" & is.na(nonA_results_stats$concordance))),
    nonA_num_gangstr_loci_passing_filter_missing_gt = length(which(nonA_results_stats$caller == "GangSTR" & is.na(nonA_results_stats$concordance))),
    nonA_num_eh_loci_passing_filter_missing_gt = length(which(nonA_results_stats$caller == "EH" & is.na(nonA_results_stats$concordance)))) %>%
  
  mutate(
    nonA_frac_num_passing_filters_over_nonA_num_loci_panel = nonA_num_loci_passing_filters / nonA_num_loci_panel,
    nonA_frac_loci_failing_filters_over_nonA_num_loci_panel = nonA_num_loci_failing_filters / nonA_num_loci_panel,
    nonA_frac_num_fully_genotyped_over_nonA_num_loci_panel = nonA_num_loci_fully_genotyped / nonA_num_loci_panel,
    nonA_frac_num_fully_genotyped_over_nonA_num_passing_filters = nonA_num_loci_fully_genotyped / nonA_num_loci_passing_filters,
    nonA_frac_num_caller_hipstr_over_nonA_num_passing_filters = nonA_num_loci_caller_hipstr / nonA_num_loci_passing_filters,
    nonA_frac_num_caller_gangstr_over_nonA_num_passing_filters = nonA_num_loci_caller_gangstr / nonA_num_loci_passing_filters,
    nonA_frac_num_caller_eh_over_nonA_num_passing_filters = nonA_num_loci_caller_eh / nonA_num_loci_passing_filters,
    nonA_frac_loci_passing_filter_missing_gt_over_nonA_num_passing_filters = nonA_num_loci_passing_filter_missing_gt / nonA_num_loci_passing_filters,
    nonA_frac_2bp_loci_passing_filter_missing_gt_over_nonA_num_2bp_passing_filters = nonA_num_2bp_loci_passing_filter_missing_gt / nonA_num_2bp_loci_passing_filters,
    nonA_frac_3bp_loci_passing_filter_missing_gt_over_nonA_num_3bp_passing_filters = nonA_num_3bp_loci_passing_filter_missing_gt / nonA_num_3bp_loci_passing_filters,
    nonA_frac_4bp_loci_passing_filter_missing_gt_over_nonA_num_4bp_passing_filters = nonA_num_4bp_loci_passing_filter_missing_gt / nonA_num_4bp_loci_passing_filters,
    nonA_frac_discordant_over_nonA_sum_concordant_discordant = nonA_num_discordant_loci / (nonA_num_concordant_loci + nonA_num_discordant_loci),
    nonA_frac_2bp_discordant_over_nonA_sum_2bp_concordant_discordant = 
           nonA_num_2bp_discordant_loci / (nonA_num_2bp_concordant_loci + nonA_num_2bp_discordant_loci),
    nonA_frac_3bp_discordant_over_nonA_sum_3bp_concordant_discordant =
           nonA_num_3bp_discordant_loci / (nonA_num_3bp_concordant_loci + nonA_num_3bp_discordant_loci),
    nonA_frac_4bp_discordant_over_nonA_sum_4bp_concordant_discordant =
           nonA_num_4bp_discordant_loci / (nonA_num_4bp_concordant_loci + nonA_num_4bp_discordant_loci),
    nonA_frac_hipstr_discordant_over_nonA_sum_hipstr_concordant_discordant =
           nonA_num_hipstr_discordant_loci / (nonA_num_hipstr_concordant_loci + nonA_num_hipstr_discordant_loci),
    nonA_frac_gangstr_discordant_over_nonA_sum_gangstr_concordant_discordant =
           nonA_num_gangstr_discordant_loci / (nonA_num_gangstr_concordant_loci + nonA_num_gangstr_discordant_loci),
    nonA_frac_eh_discordant_over_nonA_sum_eh_concordant_discordant =
           nonA_num_eh_discordant_loci / (nonA_num_eh_concordant_loci + nonA_num_eh_discordant_loci))
  

# poly A loci stats  
A_stats <-
  data.frame(
    # DELETE condition column when finished with optimization since parameters will not change
    condition = sub('.*_([0-9]+)_genotypes\\.rds$', "\\1",args[1]),
    A_num_loci_panel = nrow(A_results_stats),
    A_num_loci_passing_filters = length(which(A_results_stats$caller != "no call")),
    A_num_loci_failing_filters = length(which(A_results_stats$caller == "no call")),
    A_num_loci_caller_hipstr = length(which(A_results_stats$caller == "HipSTR" )),
    A_num_loci_caller_gangstr = length(which(A_results_stats$caller == "GangSTR" )),
    A_num_loci_caller_eh = length(which(A_results_stats$caller == "EH" )),
    A_num_loci_fully_genotyped = length(which(A_results_stats$concordance == TRUE | A_results_stats$concordance == FALSE)),
    A_num_hipstr_loci_fully_genotyped = length(which((A_results_stats$concordance == TRUE | A_results_stats$concordance == FALSE)
                                                      & A_results_stats$caller == "HipSTR" )),
    A_num_gangstr_loci_fully_genotyped = length(which((A_results_stats$concordance == TRUE | A_results_stats$concordance == FALSE)
                                                        & A_results_stats$caller == "GangSTR" )),
    A_num_eh_loci_fully_genotyped = length(which((A_results_stats$concordance == TRUE | A_results_stats$concordance == FALSE)
                                                  & A_results_stats$caller == "EH" )),
    A_num_concordant_loci = length(which(A_results_stats$concordance == TRUE )),
    A_num_hipstr_concordant_loci = length(which(A_results_stats$caller == "HipSTR" & A_results_stats$concordance == TRUE)),
    A_num_gangstr_concordant_loci = length(which(A_results_stats$caller == "GangSTR" & A_results_stats$concordance == TRUE)),
    A_num_eh_concordant_loci = length(which(A_results_stats$caller == "EH" & A_results_stats$concordance == TRUE)),
    A_num_discordant_loci = length(which(A_results_stats$concordance == FALSE )),
    A_num_hipstr_discordant_loci = length(which(A_results_stats$caller == "HipSTR" & A_results_stats$concordance == FALSE)),
    A_num_gangstr_discordant_loci = length(which(A_results_stats$caller == "GangSTR" & A_results_stats$concordance == FALSE)),
    A_num_eh_discordant_loci = length(which(A_results_stats$caller == "EH" & A_results_stats$concordance == FALSE)),
    A_num_loci_passing_filter_missing_gt = length(which(is.na(A_results_stats$concordance) & A_results_stats$caller != "no call")),
    A_num_hipstr_loci_passing_filter_missing_gt = length(which(A_results_stats$caller == "HipSTR" & is.na(A_results_stats$concordance))),
    A_num_gangstr_loci_passing_filter_missing_gt = length(which(A_results_stats$caller == "GangSTR" & is.na(A_results_stats$concordance))),
    A_num_eh_loci_passing_filter_missing_gt = length(which(A_results_stats$caller == "EH" & is.na(A_results_stats$concordance)))) %>%
  
  mutate(
    A_frac_num_passing_filters_over_A_num_loci_panel = A_num_loci_passing_filters / A_num_loci_panel,
    A_frac_loci_failing_filters_over_A_num_loci_panel = A_num_loci_failing_filters / A_num_loci_panel,
    A_frac_num_fully_genotyped_over_A_num_loci_panel = A_num_loci_fully_genotyped / A_num_loci_panel,
    A_frac_num_fully_genotyped_over_A_num_passing_filters = A_num_loci_fully_genotyped / A_num_loci_passing_filters,
    A_frac_num_caller_hipstr_over_A_num_passing_filters = A_num_loci_caller_hipstr / A_num_loci_passing_filters,
    A_frac_num_caller_gangstr_over_A_num_passing_filters = A_num_loci_caller_gangstr / A_num_loci_passing_filters,
    A_frac_num_caller_eh_over_A_num_passing_filters = A_num_loci_caller_eh / A_num_loci_passing_filters,
    A_frac_loci_passing_filter_missing_gt_over_A_num_passing_filters = A_num_loci_passing_filter_missing_gt / A_num_loci_passing_filters,
    A_frac_discordant_over_A_sum_concordant_discordant = A_num_discordant_loci / (A_num_concordant_loci + A_num_discordant_loci),
    A_frac_hipstr_discordant_over_A_sum_hipstr_concordant_discordant = 
           A_num_hipstr_discordant_loci / (A_num_hipstr_concordant_loci + A_num_hipstr_discordant_loci),
    A_frac_gangstr_discordant_over_A_sum_gangstr_concordant_discordant = 
           A_num_gangstr_discordant_loci / (A_num_gangstr_concordant_loci + A_num_gangstr_discordant_loci),
    A_frac_eh_discordant_over_A_sum_eh_concordant_discordant = 
           A_num_eh_discordant_loci / (A_num_eh_concordant_loci + A_num_eh_discordant_loci))


all_loci_stats <- cbind(nonA_stats, A_stats)

print(all_loci_stats)

# save concordance results in RDS
saveRDS(results, file = paste0(sub('\\.rds$', '', args[1]),"_",config$sample$trio,"_concordance.rds"))

# save concordance statistics in TSV file
write.table(all_loci_stats, file = paste0(sub('\\.rds$', '', args[1]),"_",config$sample$trio,"_concordance_stats.tsv"), quote = F, row.names = F)

```


# I. SINGLE PARAMETER PLOTTING

## 1. Optimization statistics
```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(gghighlight)
library(readxl)
library(patchwork)

# Excel file containing all columns from the concordance_stats.csv file generated by trio_concordance.r AND all filtering parameters
# Filtered to include only rows with single parameter comparisons
optimization_stats_r <- read_excel("optimization_stats_r.xlsx") %>% filter(condition <= 274)
```


## 2. HipSTR
### a. minimum quality
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_qual != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_qual)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.2, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.05, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_qual != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_qual)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.05, y = 98000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```

### b. minimum mean quality
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_mean_min_qual != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_mean_min_qual)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.2, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.05, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_mean_min_qual != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_mean_min_qual)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.05, y = 98000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```


### c. minimum total reads
#### i. allreads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_allreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_allreads)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 13, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_allreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_allreads)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 13, y = 107000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```

#### ii. mallreads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_mallreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_mallreads)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 12, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_mallreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_mallreads)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 12, y = 107000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```


### d. minimum mean total reads
#### i. allreads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_mean_allreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_mean_allreads, label = condition)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 12, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_mean_allreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_mean_allreads, label = condition)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 12, y = 107000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```

#### ii. mallreads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_mean_mallreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_mean_mallreads, label = condition)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 12, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_mean_mallreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_mean_mallreads, label = condition)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 12, y = 107000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```


### e. minimum allele reads
#### i. allreads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_allele_allreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_allele_allreads)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 11, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_allele_allreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_allele_allreads)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 11, y = 107000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```

#### ii. mallreads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_allele_mallreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_allele_mallreads)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 11, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_allele_mallreads != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_allele_mallreads)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 11, y = 107000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)


(discordance_plot + loci_plot)

```


### f. minimum VAF
#### i. allreads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_avaf != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_avaf)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.1, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.04, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_avaf != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_avaf)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.04, y = 98000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)


(discordance_plot + loci_plot)

```

#### ii. mallreads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_mvaf != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_mvaf)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.1, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.04, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_min_mvaf != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_min_mvaf)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.04, y = 98000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)


(discordance_plot + loci_plot)

```


### g. minimum GLDIFF
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_min_hipstr_gldiff != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_min_hipstr_gldiff)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 15, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 7, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_min_hipstr_gldiff != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_min_hipstr_gldiff)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 7, y = 107000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```

### h. max stutter fraction
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_max_hipstr_stutterfrac != 1 | condition == 1) %>%
  ggplot(aes(x = nonA_max_hipstr_stutterfrac)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.25, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.93, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_max_hipstr_stutterfrac != 1 | condition == 1) %>%
  ggplot(aes(x = nonA_max_hipstr_stutterfrac)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.93, y = 98000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```

### i. max flank indel fraction
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_max_hipstr_flindel_frac != 1 & nonA_max_hipstr_flindel_frac != .01 | condition == 1 ) %>%
  ggplot(aes(x = nonA_max_hipstr_flindel_frac)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.25, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.93, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_max_hipstr_flindel_frac != 1 | condition == 1) %>%
  ggplot(aes(x = nonA_max_hipstr_flindel_frac)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.93, y = 98000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```

### j. sample fraction
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_hipstr_sample_frac != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_sample_frac)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.25, y = 2/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.07, y = 0.0019, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_hipstr_sample_frac != 0 | condition == 1) %>%
  ggplot(aes(x = nonA_hipstr_sample_frac)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 1, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.07, y = 98000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```


## 3. GangSTR

### a. minimum quality
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_qual != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_qual)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.07, y = 0.017, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_qual != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_qual)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.07, y = 135000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```

### b. minimum mean quality
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_mean_qual != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_mean_qual)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.05, y = 0.017, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_mean_qual != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_mean_qual)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.05, y = 135000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```

### c. minimum total reads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_total_reads != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_total_reads)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 12, y = 0.017, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_total_reads != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_total_reads)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 12, y = 150000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```

### d. minimum mean total reads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_mean_total_reads != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_mean_total_reads)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 12, y = 0.017, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_mean_total_reads != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_mean_total_reads)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 12, y = 150000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```

### e. minimum allele reads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_allele_reads != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_allele_reads)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 9, y = 0.017, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_allele_reads != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_allele_reads)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 9, y = 150000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```

### f. minimum VAF
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_vaf != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_vaf)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.04, y = 0.017, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_gangstr_min_vaf != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_min_vaf)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.04, y = 150000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)


(discordance_plot + loci_plot)
```

### g. sample fraction
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_gangstr_sample_frac != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_sample_frac)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.07, y = 0.017, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_gangstr_sample_frac != 0 | condition == 5) %>%
  ggplot(aes(x = nonA_gangstr_sample_frac)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 5, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.07, y = 150000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```


## 4. Expansion Hunter

### a. minimum total reads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_eh_min_total_reads != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_min_total_reads)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 15, y = 0.0305, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_eh_min_total_reads != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_min_total_reads)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 15, y = 150000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)


(discordance_plot + loci_plot)
```

### b. minimum mean total reads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_eh_min_mean_total_reads != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_min_mean_total_reads)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 15, y = 0.033, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_eh_min_mean_total_reads != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_min_mean_total_reads)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 15, y = 150000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```

### c. minimum allele reads
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_eh_min_allele_reads != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_min_allele_reads)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 12, y = 0.033, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_eh_min_allele_reads != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_min_allele_reads)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 12, y = 150000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```

### d. min vaf
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_eh_min_vaf != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_min_vaf)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.07, y = 0.0305, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_eh_min_vaf != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_min_vaf)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.07, y = 135000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```

### e. min locus coverage
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_eh_min_lc != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_min_lc)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 8, y = 0.0335, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_eh_min_lc != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_min_lc)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 8, y = 150000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)
```

### f. sample fraction
```{r}
# Plot discordance rate for each condition plus discordance rate with no filters
discordance_plot <- optimization_stats_r %>% filter(nonA_eh_sample_frac != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_sample_frac)) +
  geom_point(aes(y = nonA_frac_discordant_over_nonA_sum_concordant_discordant)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  geom_hline(aes(yintercept = 1/10000), color = "red") +
  annotate(geom = "text", x = 0.25, y = 10/10000, label = "1 in 10,000", size = 3, color = "red") +
  annotate(geom = "text", x = 0.07, y = 0.0335, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

# Plot number of fully genotyped loci for each condition plus number of fully genotyped loci with no filters
loci_plot <- optimization_stats_r %>% filter(nonA_eh_sample_frac != 0 | condition == 3) %>%
  ggplot(aes(x = nonA_eh_sample_frac)) +
  geom_point(aes(y = nonA_num_loci_fully_genotyped)) +
  gghighlight(condition != 3, unhighlighted_params = list(color = "red", alpha = 0.4)) +
  annotate(geom = "text", x = 0.07, y = 150000, label = "no filters", size = 3, color = "red") +
  ylim(0,NA)

(discordance_plot + loci_plot)

```


# J. STUTTER PLOTS

```{r}
# no filter, HipSTR only condition
condition_1_UDP_1100 <- readRDS("coriell_udp_1_genotypes_UDP-1100_concordance.rds")
condition_1_UDP_2100 <- readRDS("coriell_udp_1_genotypes_UDP-2100_concordance.rds")
condition_1_coriell_1 <- readRDS("coriell_udp_1_genotypes_1_concordance.rds")
condition_1_coriell_2 <- readRDS("coriell_udp_1_genotypes_2_concordance.rds")
condition_1_coriell_3 <- readRDS("coriell_udp_1_genotypes_3_concordance.rds")

# combine all trios
condition_1_all <- rbind(condition_1_UDP_1100,condition_1_UDP_2100,condition_1_coriell_1,condition_1_coriell_2,condition_1_coriell_3)

rm(condition_1_UDP_1100,condition_1_UDP_2100,condition_1_coriell_1,condition_1_coriell_2,condition_1_coriell_3)
gc()


# make function for easy repeat plots
stutter_plot_func <- function(filtered_data, plot_title) {
  
  # load pre-filtered data frame (keeping only samples/loci of interest)
  plotting_data <- filtered_data %>%
    select(name, chr, motif.family, sampleID, H_GB_1, H_MALLREADS) %>%
    # split MALLREADS field so just one "bp difference|read count" unit in each column
    # automatically detects number of columns needed
    separate_wider_delim(cols = H_MALLREADS, delim = ";", names_sep = "_", too_few = "align_start") %>%
    # melt data frame so "bp difference|read count" units are in one column
    pivot_longer(cols = starts_with("H_MALLREADS"), names_to = NULL, 
                 values_to = "MALLREADS", values_drop_na = TRUE) %>%
    # separate into bp difference and read count columns
    separate_wider_delim(cols = "MALLREADS", delim = "|", names = c("bp_diff","read_count"), too_few = "align_start") %>%
    # convert bp difference and read count columns to numeric
    mutate(across(c(bp_diff, read_count), as.numeric)) %>%
    # subtract the genotype call bp difference from the MALLREADS bp difference
    # bp difference column now represents bp difference from genotype call
    mutate(bp_diff_from_call = bp_diff - H_GB_1) %>%
    # divide by motif length to get number of units different from genotype call
    mutate(unit_diff = bp_diff_from_call / nchar(motif.family))
    
  # total number of reads in MALLREADS field
  all_reads <- plotting_data %>% filter(!is.na(read_count)) %>% summarise(sum(read_count))
  # fraction of reads that do not match genotype call
  no_match <- plotting_data %>% filter(unit_diff != 0, !is.na(read_count)) %>% summarise(sum(read_count))
  # fraction of reads that have more units than the genotype call
  greater_than <- plotting_data %>% filter(unit_diff > 0, !is.na(read_count)) %>% summarise(sum(read_count))
  # fraction of reads that have fewer units than the genotype call
  less_than <- plotting_data %>% filter(unit_diff < 0, !is.na(read_count)) %>% summarise(sum(read_count))
  # fraction of reads that have n+1 units than the genotype call
  plus_one <- plotting_data %>% filter(unit_diff == 1, !is.na(read_count)) %>% summarise(sum(read_count))
  # fraction of reads that have n-1 units than the genotype call
  minus_one <- plotting_data %>% filter(unit_diff == -1, !is.na(read_count)) %>% summarise(sum(read_count))
  
  ggplot(plotting_data) +
    # make bar plot with sum of read counts at each unit difference
    geom_col(aes(x = unit_diff, y = read_count), width = 1) +
    theme_bw() +
    xlab("units different from genotype call") +
    ylab("read count") +
    labs(title = plot_title,
         # add stats to caption of plot
         caption = paste0("% not matching genotype: ", round(no_match/all_reads*100, digits = 2),
                         "\n% less than genotype: ", round(less_than/all_reads*100, digits = 2),
                         "\n% greater than genotype: ", round(greater_than/all_reads*100, digits = 2),
                         "\n% n-1: ", round(minus_one/all_reads*100, digits = 2),
                         "\n% n+1: ", round(plus_one/all_reads*100, digits = 2)))
}



condition_1_nonA <- condition_1_all %>% filter(sex == "M", chr == "chrX", motif.family != "A", H_GT != ".")
condition_1_2bp <- condition_1_all %>% filter(sex == "M", chr == "chrX", nchar(motif.family) == 2, H_GT != ".")
condition_1_3bp <- condition_1_all %>% filter(sex == "M", chr == "chrX", nchar(motif.family) == 3, H_GT != ".")
condition_1_4bp <- condition_1_all %>% filter(sex == "M", chr == "chrX", nchar(motif.family) == 4, H_GT != ".")
condition_1_A <- condition_1_all %>% filter(sex == "M", chr == "chrX", motif.family == "A", H_GT != ".")

stutter_plot_func(condition_1_nonA, "all nonA, male, chrX")
stutter_plot_func(condition_1_2bp, "2bp motifs, male, chrX")
stutter_plot_func(condition_1_3bp, "3bp motifs, male, chrX")
stutter_plot_func(condition_1_4bp, "4bp motifs, male, chrX")
stutter_plot_func(condition_1_A, "polyA, male, chrX")
```

